---
title: "Certainty tool"
author: "Clareece Nevill"
format: docx
editor: visual
bibliography: references-certainty.bib
csl: american-medical-association.csl
number-sections: true
crossref:
  labels: arabic        
  subref-labels: alpha a
  chapters: true
---

This document is to go over the details of the certainty tool part of the update framework.

# Datasets

For various elements of the certainty tool, datasets from existing living systematic reviews were needed to help define default threshold values. Based on the work from work package 1, we chose the 12 Cochrane living systematic reviews that were identified by the survey that used GRADE, were in a clinical field, included a pairwise meta-analysis, and more than 1 study/RCT. We are sticking with Cochrane as they have a unified data extraction system, which will simplify data collection. Multiple outcomes were extracted to gain more data (however, only outcomes measured on a risk ratio or mean difference scale were included for simplicity). For each outcome, the meta-analysis study data was extracted as well as meta-analysis model details and how many levels the evidence was downgraded for each GRADE domain (i.e., 0, 1, or 2).

## [Antibiotics COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015025/full "Cochrane website")

Below outcomes are from October 2021 version regarding inpatients with confirmed moderate to severe COVID-19.

-   All-cause mortality at day 28 based on 4 RCTs.

-   Improvement of clinical status at day 28, based on 3 RCTs.

-   Serious adverse events, based on 4 RCTs.

-   Any adverse events, based on 3 RCTs.

-   Cardiac arrhythmias, based on 4 RCTs.

## [Chronic rhinosinusitis](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD013513.pub3/full "Cochrane website")

Below are outcomes from the March 2021 version (as studies were the same in earlier versions), for the dupilimab versus placebo comparison.

-   Health-related quality of life - disease-specific (SNOT-22, range 0 to 110, lower = better) based on 3 studies.

-   Disease severity - VAS (range 0 to 10, lower = better) based on 3 studies.

-   Serious adverse events based on 3 studies.

-   Avoidance of surgery based on 2 studies.

-   Adverse events based on 3 studies.

## [Colchicine COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015045/full "Cochrane website")

Below outcomes are for hospitalised patients with COVID-19 and moderate to severe disease. Number of studies are as of October 2021.

-   All-cause mortality at day 28 based on 2 studies.

-   Worsening of clinical status up to day 28 based on 2 studies.

## [Convalescent plasma COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD013600.pub6/full "Cochrane website")

Below outcomes are for convalescent plasma vs placebo or standard care in those with moderate to severe disease.

-   All-cause mortality at up to day 28. In the May 2023 (and February 2023) version, this estimate was based on 21 RCTs. In the May 2021 version, the estimate was based on 7 RCTs. In earlier versions, the outcomes were not the exact same.

-   Clinical worsening at up to day 28. In the May 2023 (and February 2023) version, this was based on 6 RCTs. In earlier versions, the outcomes were not the exact same.

-   Clinical improvement at up to day 28. In the May 2023 (and February 2023) version, this was based on 6 RCTs. In earlier versions, the outcomes were not the exact same.

-   Grades 3 and 4 adverse events up to day 28. In May 2023 (and February 2023), this was based on 6 RCTs. In May 2021, this was based on 4 RCTs. In earlier versions, the outcomes were not the exact same. When looking at the May/February 2023 versions, it is not clear which studies were included; in the text it mentions the study Avendano-Sola 2021, but in the forest plots it mentions AlQahtani 2021. For this reason, we are not including these results.

-   Serious adverse events up to day 28. In May 2023 (and February 2023), this was based on 6 RCTs. In earlier versions, the outcomes were not the exact same.

## [Fluvoxamine COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015391/full "Cochrane website")

The below outcomes are for September 2022 version.

-   All-cause mortality at day 28 based on 2 RCTs.

-   All-cause admission to hospital or death (before hospital admission) based on 2 RCTs.

-   Serious adverse events based on 2 RCTs.

-   Any adverse events based on 2 RCTs.

## [Inhaled corticosteroids COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015125/full "Cochrane website")

The below outcomes are from March 2022.

-   All-cause mortality (up to 30 days) based on 3 studies.

-   Admission to hospital or death (up to 30 days) based on 2 studies.

-   Symptom resolution (all initial symptoms resolves at day 14) based on 2 studies.

## [Ivermectin COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015017.pub3 "Cochrane website")

Below outcomes are for those in inpatient settings with moderate to severe disease.

-   All-cause mortality at day 28. In the June 2022 version, the estimate was based on 3 RCTs. In the July 2021 version, it was based on 2 RCTs.

-   Serious adverse events. In the June 2022 version, this was based on 2 RCTs. This outcome was not present in earlier versions.

-   Any adverse events. In June 2022, this was based on 3 RCTs. In July 2021, this was based on 1 RCT.

-   Viral clearance at 7 days. In June 2022, this was based on 3 RCTS. In earlier versions, the outcome was not the exact same definition.

## [Janus kinase inhibitors COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015209/full "Cochrane website")

Outcomes for June 2022 version.

-   All-cause mortality (up to day 28) based on 6 RCTs.

-   All-cause mortality (up to day 60) based on 2 RCTs.

-   Adverse events (any grade) based on 3 RCTs.

-   Serious adverse events based on 4 RCTs.

-   Secondary infection based on 4 RCTs.

## [Non-small cell lung cancer](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD013257.pub3 "Cochrane website")

-   Outcome of adverse events of grade 3-4 (for same comparison above). In April 2021 (and December 2020), this was based on 5 studies.

## [Oral anticoagulation](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD006466.pub7/full "Cochrane website")

Looking at VKA compared to no prophylaxis

-   Mortality at 6 months. As of October 2021, this was based on 3 RCTs.

-   Mortality at 12 months. As of October 2021 (and December 2017), this was based on 5 RCTs.

-   Major bleeding (follow-up 12 months). As of October 2021 (and December 2017), this was based on 5 RCTs.

-   Minor bleeding (follow-up 12 months). As of October 2021 (and December 2017), this was based on 4 RCTs.

## [Remdesivir COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD014962.pub2/full "Cochrane website")

Below outcomes are for individuals with moderate to severe COVID-19.

-   All-cause mortality up to day 28. In January 2023 (and August 2021), this was based on 4 RCTs.

-   Adverse events (any grade) up to day 28. In January 2023, this was based on 4 RCTs. In August 2021, this was based on 3 RCTs.

-   Serious adverse events up to day 28. In January 2023, this was based on 4 RCTs. In August 2021, this was based on 3 RCTs.

## [Systemic corticosteroids COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD014963.pub2/full "Cochrane website")

Below outcomes are for corticosteroids plus standard care vs standard care (plus/minus placebo).

-   All-cause mortality up to day 30. In November 2022, this was based on 10 RCTs. The outcome for August 2021 was split differently, and so can't compare.

-   All-cause mortality up to day 120. In November 2022, this was based on 3 RCTs. The outcome for August 2021 was split differently, and so can't compare.

# Risk of Bias

The risk of bias data should ideally have been completed by the time researchers use the update framework. Therefore, the risk of bias element of the certainty tool is about deciding:

-   What proportions of studies were categorised at different levels of risk

-   Whether any plausible or crucial bias that has been identified in studies is likely to seriously alter the results.

These two key factors - defined by Cochrane - decide whether to downgrade the results, and by how many levels. However, the Cochrane advice doesn't give specifics to these factors (e.g., uses phrases such as 'most information', rather than a specific proportion).

## Cochrane guidance

### Proportion of information

The three broad categories are:

1.  Most information is from studies at low risk of bias

2.  Most information is from studies at low or unclear risk of bias

3.  Proportion of information from studies at high risk of bias is sufficient to affect the interpretation of results

### Likely to alter results

Within category (2) above, it is further split into:

-   Plausible bias will **unlikely** seriously alter the results -\> no downgrading

-   Plausible bias will **likely** seriously alter the results -\> downgrade one level

Within category (3) above, it is further split into:

-   Crucial risk likely to **seriously** alter the results -\> downgrade one level

-   Crucial risk likely to **very seriously** alter the results -\> downgrade two levels

## Risk of Bias tools

There are differences between ROB1 and RoB2 (which both feature in the datasets). Whilst RoB 2 is the later version and thus recommended, it is a lot more time consuming - as such, many researchers still stick with RoB 1. Therefore, it would be ideal if the certainty tool could (technically) function regardless of which RoB tool was used.

In the Cochrane data/reports, when RoB 1 tool was used an overall rating doesn't appear to be calculated. When this is the case, we can use Stewart et al's probability matrices to give an overall rating that can be used (as RoB 2 gives an overall rating). Note - in the overall probability matrix for risk of bias, for some reason there aren't rows for when 'Other' could be classified as 'unclear'. Therefore, in such cases, we've made the assumption that 'unclear' holds the same significance as if it was rated as 'high' (especially given that Stewart's network appeared to be more 'optimistic' than Cochrane, and researchers have indicated that they wish for more false positives, than false negatives).

::: callout-note
## Different results

The developed tool may behave differently depending on the Risk of Bias tool that was used. Often, the ROB1 tool was used across all outcomes (rather than individually), and the overall rating for each study is having to be calculated. Therefore, the resulting overall RoBs may be missing important factors that are only specific to certain outcomes (see where chronic rhinosinusitis example is looked at further in section @sec-downgraded-1-level-but-score-of-less-than-1.5). Either the tool only gets used on ROB2 datasets, or it comes with a warning if using ROB1, and maybe the 'adjustment factor' idea can 'fix' these cases.
:::

## Design of RoB element

I've created a function that calculates the weighted average risk of bias, based on the meta-analytic weights used and the overall risk of bias for each study (with respect to the specific outcome where possible). Calculating a weighted average may take into account both of the guidance elements mentioned above (as a high risk study that gives a small weight is unlikely to change the results (depending on how different the estimate is from the pooled result)).

Let's see what the weighted average RoB is for each of our datasets.

```{r, echo = FALSE, output = FALSE}
library(dplyr)
library(purrr)
library(knitr)
library(kableExtra)
source("RoB_downgrades.R")
```

```{r, echo = FALSE}

# Read in info from all datasets
primary_data <- readxl::read_excel("Datasets/primary_info.xlsx", sheet = "Studies")
GRADE_results <- readxl::read_excel("Datasets/primary_info.xlsx", sheet = "GRADE")
primary_data$Weight[primary_data$Weight == "NA" | is.na(primary_data$Weight)] <- "0"  # Where there is no weight associated, specify it as zero.
primary_data$`other bias`[primary_data$`other bias` == "Unclear risk"] <- "High risk"

# Calculate average for each dataset and add to GRADE dataframe
GRADE_results$RoB_avg <- mapply(function(review_type, version_date, outcome, Rob_tool_used) {
  # subset primary_data
  data <- primary_data %>% 
    filter(Review == review_type & as.Date(Added) <= as.Date(version_date) & Outcome == outcome)
  # run RoB_downgrades()
  result <- weighted_RoB(
    data = data,
    rob_tool = Rob_tool_used,
    overall_rob = "Overall bias",
    random_selection = "random sequence generation (selection bias)",
    allocation_selection = "allocation concealment (selection bias)",
    performance = "performance bias",
    detection = "detection bias",
    attrition = "attrition bias",
    reporting = "reporting bias",
    other = "other bias",
    weights = "Weight"
  )
  # return average grade
  result
},
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome, GRADE_results$Rob_tool)

# Display results
knitr::kable(GRADE_results %>% select(c("Review", "Version", "Outcome", "Rob_tool", "Downgrade RoB", "RoB_avg")) %>% arrange(RoB_avg), digits = 2)
```

From the above, one can see that at a weighted average RoB score of 1.5, the reviews started getting downgraded by 1 level. None of the example datasets downgraded by two levels for risk of bias, yet some did have the maximum weighted average RoB score of 3.

::: callout-important
## Proposed threshold

Based on the current data, a basic default threshold would be that when the weighted average RoB score is:

-   Greater than or equal to 1.5 - downgrade by one level

-   When to downgrade two levels is less simple and explored further below
:::

In the next sections, I'm going to investigate a bit further the studies that didn't quite follow the general pattern to see if there is anything else quantifiable that can be used to 'fine-tune' the tool.

### Downgraded 1 level, but score of less than 1.5 {#sec-downgraded-1-level-but-score-of-less-than-1.5}

In the chronic rhinosinusitis dataset, for outcomes VAS (disease severity) and avoidance of surgery, the weighted average RoB score was 1 (the lowest possible), yet the authors downgraded the evidence by one level for risk of bias. Note that they used the RoB 1 tool, and assessed it across all outcomes.

For VAS, this was due to "s*tudy limitations: methods or criteria used in the measurement of the outcome were not validated*". In the risk of bias table that is available, the only related information I can find is within the 'selective reporting' field, that there was some lack of clarity regarding the choice of measurement tool, and that for VAS specifically, the outcome was not pre-specified in the original trial registry data. This led to an overall rating of 'Unclear risk' for all 3 studies (note, across all outcomes).

For avoidance of surgery, this was due to "*serious limitations: the criteria used for requiring/not requiring 'rescue surgery' were unclear*". This is similar to the above outcome, in that 'lack of clarity' is mentioned. Specifically, the number of participants who required surgery was missing. As above, this led to an overall rating of 'Unclear' risk for all 3 studies.

As the researchers didn't give an overall risk of bias, my function had to estimate it, and it put all 3 studies as most likely being 'low risk' (note, with 100% probability according to Stewart et. al.'s network).

#### What does this mean?

The big problem here is that the RoB 1 tool doesn't distinguish between different outcomes, and just considers it all together, yet GRADE wants researchers to consider each selection of evidence regarding each specific outcome. This is essentially incompatible.

::: callout-important
## RoB 1 caution

My suggestion is therefore to say that this tool will not perform as well if RoB 1 is being used, due to the above. Researchers can use the proposed tool, but with caution, and will likely have to use the adjustment factor proposed in the below section @sec-adjustment-factor.
:::

### Didn't downgrade, but score of more than 1.5

In the convalescent plasma dataset, for serious adverse events a weighted average RoB score of 1.79 was given for the Feburary 2023 version (which was the same studies in May 2023). In the oral anticoagulation dataset, for outcomes mortality at 6 months and 12 months, a weighted average RoB score of 2.88 and 2.66 was given respectively. Yet for these three outcomes, the researchers did not downgrade the evidence regarding risk of bias.

For serious adverse events in the convalescent plasma dataset, most of the studies were of 'some concerns'. The RoB score for serious adverse events of 1.79 not matching up with downgrading isn't of huge concern and could be categorised as being down to 'difference of opinion' on their significance.

For the mortality outcomes in the oral anticoagulation dataset, the researchers stated that there was "*some concern with risk of bias associated with unclear allocation concealment, taken into account when downgrading from high to moderate certainty*". For both outcomes, the GRADE rating was 'Moderate', downgraded by one level due to serious imprecision. It would seem that the researchers felt this downgrading incorporated the concerns regarding risk of bias, and so no further downgrading was required.

#### What does this mean?

Both scenarios could be corrected with the adjustment factor being proposed below. For convalescent plasma, the 'difference of opinions' between the default settings and the researchers could be adjusted by changing the threshold (going from 1.5 to 1.8 wouldn't be a huge adjustment). For oral coagulation, the researchers are clearly doing their own thing regarding risk of bias, which they can account for by changing the threshold for downgrading (again) due to risk of bias to a much higher value.

### Downgraded only 1 level, but score of 3

There were five instances where all studies have overall high risk of bias, yet the review team only downgraded the evidence by one level due to risk of bias (Colchicine COVID-19 - mortality and worsening; inhaled corticosteroids COVID-19 - symptom resolution; fluvoxamine COVID-19 - serious adverse effects and any adverse effects). The guidelines for downgrading for two levels has the additional factor than any crucial risk identified needs to also likely very seriously alter the results. The standard way to assess this would be to conduct a sensitivity analysis, removing the studies that had high risk of bias. However, in the case where all studies were deemed to be high risk (such as these five examples), this would result in no effect estimate being available, and therefore, no comparison available.

#### What does this mean?

I get the feeling that is is *very* rare for the evidence to be downgraded by two levels due to risk of bias. This is from experience from extracting data (found no examples), and the fact that where there is a high amount of risk of bias, a sensitivity analysis may not be informative (as described above).

::: callout-important
## Default second threshold

Therefore, I think it would be useful to have an option in the tool to state that there is **only one threshold** (i.e., for downgrading one level), and for that option to be the default. If reviewers wish to, they can change this themselves and add a second threshold (that they define) for when to downgrade two levels.
:::

### Adjustment factor {#sec-adjustment-factor}

So far, the idea of an adjustment factor seems sensible as, with the current approach, some research teams downgrade based on risk of bias due to different thresholds/criteria (see sections above); the data collected did not give a straight relationship with the weighted average RoB.

A first approach to this, that is very rudimentary, is to move the default threshold such that the weighted average RoB 'correctly' matches to the set levels the researchers downgraded the outcome for in the most recent version. The new thresholds are then used when considering the review with new studies. This can be thought of as 'calibrating' the tool to the review, and so it'll be focusing on the [likely change]{.underline} from the addition of new studies, rather than a change due to the differences between researchers judgement and tool defaults.

This has been added to the developed function(s) regarding risk of bias.

# Imprecision

The following are suggested ways to quantitatively assess imprecision:

-   Are there enough events?

-   Is the sample size large enough?

-   Does the confidence interval cover multiple messages?

Let's consider each one in a bit more detail regarding practically incorporating it into the tool.

## Are there enough events?

Based on what the smoking review does, they have the following rule based on the total number of events across all arms and all studies:

-   If there are \< 100 events - downgrade 2 levels

-   If there are \< 300 events - downgrade 1 level

However, one does need to consider rare events, and so these thresholds will likely have to be editable. Furthermore, it is only suitable for reviews of dichotomous outcomes. Otherwise, this is a pretty easy method to implement.

## Does the confidence interval cover multiple messages?

This approach is a very practical way of assessing imprecision by actively seeing if the review is suggesting multiple 'messages' (i.e., clinical harm and benefit), a direct consequence of imprecision. It does so by the following categories:

-   If the confidence interval (CI) includes a meaningfully positive or negative effect, as well as no meaningful effect in the opposite direction, then downgrade 1 level

-   If the CI includes both meaningfully positive and negative effects, then downgrade 2 levels

The main element that needs defining in this approach is what value translates to a 'meaningful' effect? The following defaults have been suggested by the literature:

-   For dichotomous outcomes, a meaningful effect is when the relative risk reduction or increase is greater than 25%

-   For continuous outcomes, the minimal importance difference should be used if known, otherwise a default of +/- 0.5 can be used.

Again, as with considering number of events, these thresholds need to be editable by the reviewers where appropriate. This includes where the review topic is considering rare events.

## Is the sample size large enough?

This has been suggested by multiple sources. For dichotomous outcomes, it has been suggested to consider whether the total sample size across all trials included in the review is lower than an optimal information size. If so, downgrade 1 level. However, how should the optimal information size be calculated? It is often suggested to calculate the value as if it is a single trial and compare, but there are equations out there for calculating the optimal number of trials and average trial sample size for a meta-analysis. For continuous outcomes, a 'rule of thumb' that has been suggested is that is there are \< 400 participants in total across the review, then downgrade 1 level.

Due to the discrepancies around how to calculate an optimal information size, I'm tempted to not utilise this approach. Additionally, I believe the other two approaches above probably cover any imprecision that would be 'found' by considering the total sample size of the review.

## Testing tool

We have two approaches for downgrading, both of which can lead to downgrading 2 levels. However, the tool can only downgrade a maximum of 2 levels (as per GRADE guidance). Where both approaches are appropriate (i.e., for dichotomous outcomes), the minimum number of levels downgraded will be taken. 'Summing' them doesn't make sense as both approaches are attempting to 'measure' the same thing, and so they shouldn't accumulate. Taking the minimum was preferred over taking the maximum as one approach may highlight precision, but if taking the maximum, the other approach may downgrade it unecessarily (e.g., if the confidence interval is narrow and only inclusing one 'interpretation', but the number of the events is slightly small). Furthermore, when taking the minimum rather than the maximum, this married up better with our datasets (agreement in 72% of cases compared with 57%). In addition, I also tested to see whether just sticking with the confidence interval approach improved agreement, but it didn't (agreement of 66%).

Let's see how well the tool performs at predicting the imprecision grading levels using the default values suggested by the literature.

```{r, echo = FALSE, output = FALSE}
source("Imprecision_downgrades.R")
```

```{r, echo = FALSE}
# Calculate total events and extract confidence intervals for each dataset and add to GRADE dataframe
GRADE_results$total_events <- mapply(function(review_type, version_date, outcome, measure) {
  if (measure == "RR") {
    # subset primary_data
    data <- primary_data %>% 
      filter(Review == review_type & as.Date(Added) <= as.Date(version_date) & Outcome == outcome)
    # run sum_events()
    sum_events(
      data = data,
      event_cols = c("Experimental cases", "Control cases")
    )
  } else {
    NA
  }
},
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome, GRADE_results$Measure)

GRADE_results$CI <- mapply(function(review_type, version_date, outcome, measure, model) {
  # subset primary_data
  data <- primary_data %>% 
    filter(Review == review_type & as.Date(Added) <= as.Date(version_date) & Outcome == outcome)
  # run meta-analysis
  if (measure == 'RR') {
    rma_df <- metafor::escalc(data = data, ai = as.numeric(`Experimental cases`), ci = as.numeric(`Control cases`), n1i = `Experimental N`, n2i = `Control N`, measure = measure)
  } else if (measure == 'MD') {
    rma_df <- metafor::escalc(data = data, n1i = `Experimental N`, n2i = `Control N`, m1i = as.numeric(`Experimental mean`), m2i = as.numeric(`Control mean`), sd1i = as.numeric(`Experimental SD`), sd2i = as.numeric(`Control SD`), measure = measure)
  }
  meta <- metafor::rma(data = rma_df, yi = yi, vi = vi, measure = measure, method = model)
  if (measure == 'RR') {
    paste0(round(exp(meta$ci.lb), 2), ", ", round(exp(meta$ci.ub), 2))
  } else {
    paste0(round(meta$ci.lb, 2), ", ", round(meta$ci.ub, 2))
  }
}
,
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome, GRADE_results$Measure, GRADE_results$Model)

# Predict downgrading levels based on the information
GRADE_results$imprecision_levels <- mapply(function(review_type, version_date, outcome, measure, total_events, CI) {
  
  # create interpretation dataframe
  if (measure == "RR") {
    interpretation_df <- CI_interpretation_bin(
      CI_lb = as.numeric(stringr::str_split_fixed(CI, ",", n = 2)[1]),
      CI_ub = as.numeric(stringr::str_split_fixed(CI, ",", n = 2)[2])
    )
  } else if (measure == "MD") {
    interpretation_df <- CI_interpretation_cont(
      CI_lb = as.numeric(stringr::str_split_fixed(CI, ",", n = 2)[1]),
      CI_ub = as.numeric(stringr::str_split_fixed(CI, ",", n = 2)[2])
    )
  }
  
  # predict number of levels to downgrade by
  imprecision_downgrades(
    total_events = total_events, interpretation_df = interpretation_df
  )

},
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome, GRADE_results$Measure, GRADE_results$total_events, GRADE_results$CI)

# Display results
knitr::kable(GRADE_results %>% select(c("Review", "Version", "Outcome", "Downgrade Imprecision", "total_events", "CI", "imprecision_levels")) %>% arrange(total_events), digits = 2)
```

## Exploring differences

Of the 47 outcomes, the tools prediction of the number of levels to downgrade matched the reviewers assessment for `` sum(GRADE_results$`Downgrade Imprecision` == GRADE_results$imprecision_levels) `` outcomes. Let's explore those that didn't match and see if we can know the reason.

### Instances where tool over-downgraded

There were 4 cases where the tool downgrading more levels for imprecision than the reviewers did. This happened for the following outcomes:

-   Convalescent plasma - Adverse events

    -   As of 22-Oct-25, the review is locked, so can't see details on why only downgraded 1 level. But there were only 69 events, and a CI of 0.58 to 1.41 which is quite wide.

-   Ivermectin COVID-19 - Adverse events

    -   As of 22-Oct-25, the review is locked, so can't see details on why only downgraded 1 level. But there were 47 events, and a CI of 0.61 to 1.79, which is again quite a wide interval.

-   Janus kinase inhibitors - Mortality at day 60

    -   There were 249 events and a CI of 0.56 to 0.86. The reviewers didn't downgrade the evidence at all for imprecision, which I can see. The number of events is only a little less than 300, and the confidence interval is quite well situated to 'one side', however, the upper bound was still above the threshold of 0.75. In this case, I would be tempted to say that the reviewers would consider a different threshold for defining a 'meaningful effect'.

-   Antibiotics COVID-19 - Serious adverse events

    -   There were 185 events and a CI of 0.89 to 1.4, but the reviewers didn't downgrade the evidence at all for imprecision. Considering the outcome is serious adverse events, having 185 events may not be considered as a small amount. Furthermore, the reviewers may not consider an upper bound of 1.4 to mean there are results in the meaningful positive effect region.

::: callout-tip
## Adjustment factor proposal

Adding an adjustment factor here is less simple than with risk of bias as we have two approaches for dichotomous outcomes.

If the outcome is likely to be a rare event (e.g., serious adverse events) - and indicated by the reviewers as so - then adjust the threshold regarding total events so that a previous GRADE result marries up. Otherwise, adjust the thresholds related to what a 'meaningful' effect is for assessing the the placement of the confidence interval.
:::

::: callout-caution
## Adjustment factor warning

Currently, I've put in code that allows the threshold to automatically change if the tool didn't correctly predict the downgrading levels on a previous version of the review. However, if reviewers are competent and have a suitable system for carrying out GRADE, then they should be able to input their own thresholds that should allow the tool to marry up with previous versions.

As such, the tool/code should only have the automatic 'threshold adjuster' if it is specifically asked for. These cases would be for myself to test it on reviews for which I don't have contact with the authors, and potentially reviews that are midway through being living, and wish to have some help with defining the thresholds. Therefore, the tool should also output which thresholds it used (so that users can be confident in knowing what was used).
:::

### Instances where the tool under-graded

There were 9 cases where the tool downgrading less levels for imprecision than the reviewers did. This happened for the following outcomes:

-   Fluvoxamine COVID-19

    -   Adverse events

        -   There were 201 events and a CI of 0.82 to 1.37, yet the reviewers downgraded it 2 levels for imprecision.

    -   Hospitalisation or death

        -   There were 181 events and a CI of 0.16 to 1.88, yet the reviewers downgraded it 2 levels for imprecision. If the tool took the maximum levels rather than the minimum, then the tool would have agreed as the CI does include meaningful benefit and harm (i.e., it is very wide).

    -   Serious adverse events

-   Oral anticoagulation

    -   Mortality at 6 months

    -   Mortality at 12 months

-   Remdesivir COVID-19

    -   Mortality

    -   Adverse events (for Jan '23 and Aug '21 update)

    -   Serious adverse events

# Indirectness

There are two elements of indirectness that can occur: a difference between studies included and review aims, and when indirect comparisons are utilised due to lack of direct evidence[@guyatt2025]. As we're only considering pairwise meta-analyses and RCTs, for now we're going to assume that the reviewer has not needed to find indirect evidence, and so we will focus on the former element of indirectness. Otherwise, the GRADE guidance for authors suggests considering whether the evidence found is more restrictive than the review question in terms of PICO[@schünemann2013]. Due to the nature of this domain, there will be no quantitative elements, and instead, the tool will have the following prompt questions from the user to answer.

-   Are the populations presented in the current evidence so much more restrictive than the target population of the review that they're likely to affect the magnitude of effect substantially? If so, is it serious or very serious?

-   Are the interventions that have been studied in the current pool of evidence more restrictive than the target interventions of the review? If so, is it serious or very serious?

-   Are the comparators that have been included in the current studies more restrictive than the target comparator(s) of the review? If so, is it serious or very serious?

-   Are the outcomes that are presented in the current evidence more restrictive than the target outcomes of the review? If so, is it serious or very serious?

The questions are not to be answered on their own but as a collective as the judgement of whether to downgrade due to indirectness is to be done based on the combined effect of the areas considered above[@guyatt2011]. As such, the user will simply be asked to judge the evidence based on the prompt questions, as to whether any problems identified need to be rates down by two, one, or no levels. Additionally, when presented with the proposed tool, reviewers will also be given links to guidelines written by GRADE developers[@guyatt2011a; @guyatt2025a]

Unfortunately, due to the nature of this domain, I cannot 'test' it, or measure it against any of the reviews I've collated.

# References

::: {#refs}
:::
