---
title: "Certainty tool"
author: "Clareece Nevill"
engine: knitr
format: docx
editor: visual
bibliography: references-certainty.bib
csl: american-medical-association.csl
number-sections: true
crossref:
  labels: arabic        
  subref-labels: alpha a
  chapters: true
---

This document is to go over the details of the certainty tool part of the update framework.

# Datasets

For various elements of the certainty tool, datasets from existing living systematic reviews were needed to help define default threshold values. Based on the work from work package 1, we chose the 12 Cochrane living systematic reviews that were identified by the survey that used GRADE, were in a clinical field, included a pairwise meta-analysis, and more than 1 study/RCT. We are sticking with Cochrane as they have a unified data extraction system, which will simplify data collection. Multiple outcomes were extracted to gain more data (however, only outcomes measured on a risk ratio or mean difference scale were included for simplicity). For each outcome, the meta-analysis study data was extracted as well as meta-analysis model details and how many levels the evidence was downgraded for each GRADE domain (i.e., 0, 1, or 2).

## [Antibiotics COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015025/full "Cochrane website")

Below outcomes are from October 2021 version regarding inpatients with confirmed moderate to severe COVID-19.

-   All-cause mortality at day 28 based on 4 RCTs.

-   Improvement of clinical status at day 28, based on 3 RCTs.

-   Serious adverse events, based on 4 RCTs.

-   Any adverse events, based on 3 RCTs.

-   Cardiac arrhythmias, based on 4 RCTs.

## [Chronic rhinosinusitis](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD013513.pub3/full "Cochrane website")

Below are outcomes from the March 2021 version (as studies were the same in earlier versions), for the dupilimab versus placebo comparison.

-   Health-related quality of life - disease-specific (SNOT-22, range 0 to 110, lower = better) based on 3 studies.

-   Disease severity - VAS (range 0 to 10, lower = better) based on 3 studies.

-   Serious adverse events based on 3 studies.

-   Avoidance of surgery based on 2 studies.

-   Adverse events based on 3 studies.

## [Colchicine COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015045/full "Cochrane website")

Below outcomes are for hospitalised patients with COVID-19 and moderate to severe disease. Number of studies are as of October 2021.

-   All-cause mortality at day 28 based on 2 studies.

-   Worsening of clinical status up to day 28 based on 2 studies.

## [Convalescent plasma COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD013600.pub6/full "Cochrane website")

Below outcomes are for convalescent plasma vs placebo or standard care in those with moderate to severe disease.

-   All-cause mortality at up to day 28. In the May 2023 (and February 2023) version, this estimate was based on 21 RCTs. In the May 2021 version, the estimate was based on 7 RCTs. In earlier versions, the outcomes were not the exact same.

-   Clinical worsening at up to day 28. In the May 2023 (and February 2023) version, this was based on 6 RCTs. In earlier versions, the outcomes were not the exact same.

-   Clinical improvement at up to day 28. In the May 2023 (and February 2023) version, this was based on 6 RCTs. In earlier versions, the outcomes were not the exact same.

-   Grades 3 and 4 adverse events up to day 28. In May 2023 (and February 2023), this was based on 6 RCTs. In May 2021, this was based on 4 RCTs. In earlier versions, the outcomes were not the exact same. When looking at the May/February 2023 versions, it is not clear which studies were included; in the text it mentions the study Avendano-Sola 2021, but in the forest plots it mentions AlQahtani 2021. For this reason, we are not including these results.

-   Serious adverse events up to day 28. In May 2023 (and February 2023), this was based on 6 RCTs. In earlier versions, the outcomes were not the exact same.

## [Fluvoxamine COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015391/full "Cochrane website")

The below outcomes are for September 2022 version.

-   All-cause mortality at day 28 based on 2 RCTs.

-   All-cause admission to hospital or death (before hospital admission) based on 2 RCTs.

-   Serious adverse events based on 2 RCTs.

-   Any adverse events based on 2 RCTs.

## [Inhaled corticosteroids COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015125/full "Cochrane website")

The below outcomes are from March 2022.

-   All-cause mortality (up to 30 days) based on 3 studies.

-   Admission to hospital or death (up to 30 days) based on 2 studies.

-   Symptom resolution (all initial symptoms resolves at day 14) based on 2 studies.

## [Ivermectin COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015017.pub3 "Cochrane website")

Below outcomes are for those in inpatient settings with moderate to severe disease.

-   All-cause mortality at day 28. In the June 2022 version, the estimate was based on 3 RCTs. In the July 2021 version, it was based on 2 RCTs.

-   Serious adverse events. In the June 2022 version, this was based on 2 RCTs. This outcome was not present in earlier versions.

-   Any adverse events. In June 2022, this was based on 3 RCTs. In July 2021, this was based on 1 RCT.

-   Viral clearance at 7 days. In June 2022, this was based on 3 RCTS. In earlier versions, the outcome was not the exact same definition.

## [Janus kinase inhibitors COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD015209/full "Cochrane website")

Outcomes for June 2022 version.

-   All-cause mortality (up to day 28) based on 6 RCTs.

-   All-cause mortality (up to day 60) based on 2 RCTs.

-   Adverse events (any grade) based on 3 RCTs.

-   Serious adverse events based on 4 RCTs.

-   Secondary infection based on 4 RCTs.

## [Non-small cell lung cancer](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD013257.pub3 "Cochrane website")

-   Outcome of adverse events of grade 3-4 (for same comparison above). In April 2021 (and December 2020), this was based on 5 studies.

## [Oral anticoagulation](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD006466.pub7/full "Cochrane website")

Looking at VKA compared to no prophylaxis

-   Mortality at 6 months. As of October 2021, this was based on 3 RCTs.

-   Mortality at 12 months. As of October 2021 (and December 2017), this was based on 5 RCTs.

-   Major bleeding (follow-up 12 months). As of October 2021 (and December 2017), this was based on 5 RCTs.

-   Minor bleeding (follow-up 12 months). As of October 2021 (and December 2017), this was based on 4 RCTs.

## [Remdesivir COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD014962.pub2/full "Cochrane website")

Below outcomes are for individuals with moderate to severe COVID-19.

-   All-cause mortality up to day 28. In January 2023 (and August 2021), this was based on 4 RCTs.

-   Adverse events (any grade) up to day 28. In January 2023, this was based on 4 RCTs. In August 2021, this was based on 3 RCTs.

-   Serious adverse events up to day 28. In January 2023, this was based on 4 RCTs. In August 2021, this was based on 3 RCTs.

## [Systemic corticosteroids COVID-19](https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD014963.pub2/full "Cochrane website")

Below outcomes are for corticosteroids plus standard care vs standard care (plus/minus placebo).

-   All-cause mortality up to day 30. In November 2022, this was based on 10 RCTs. The outcome for August 2021 was split differently, and so can't compare.

-   All-cause mortality up to day 120. In November 2022, this was based on 3 RCTs. The outcome for August 2021 was split differently, and so can't compare.

# Risk of Bias

The risk of bias data should ideally have been completed by the time researchers use the update framework. Therefore, the risk of bias element of the certainty tool is about deciding:

-   What proportions of studies were categorised at different levels of risk

-   Whether any plausible or crucial bias that has been identified in studies is likely to seriously alter the results.

These two key factors - defined by Cochrane - decide whether to downgrade the results, and by how many levels. However, the Cochrane advice doesn't give specifics to these factors (e.g., uses phrases such as 'most information', rather than a specific proportion).

## Cochrane guidance

### Proportion of information

The three broad categories are:

1.  Most information is from studies at low risk of bias

2.  Most information is from studies at low or unclear risk of bias

3.  Proportion of information from studies at high risk of bias is sufficient to affect the interpretation of results

### Likely to alter results

Within category (2) above, it is further split into:

-   Plausible bias will **unlikely** seriously alter the results -\> no downgrading

-   Plausible bias will **likely** seriously alter the results -\> downgrade one level

Within category (3) above, it is further split into:

-   Crucial risk likely to **seriously** alter the results -\> downgrade one level

-   Crucial risk likely to **very seriously** alter the results -\> downgrade two levels

## Risk of Bias tools

There are differences between ROB1 and RoB2 (which both feature in the datasets). Whilst RoB 2 is the later version and thus recommended, it is a lot more time consuming - as such, many researchers still stick with RoB 1. Therefore, it would be ideal if the certainty tool could (technically) function regardless of which RoB tool was used.

In the Cochrane data/reports, when RoB 1 tool was used an overall rating doesn't appear to be calculated. When this is the case, we can use Stewart et al's probability matrices to give an overall rating that can be used (as RoB 2 gives an overall rating). Note - in the overall probability matrix for risk of bias, for some reason there aren't rows for when 'Other' could be classified as 'unclear'. Therefore, in such cases, we've made the assumption that 'unclear' holds the same significance as if it was rated as 'high' (especially given that Stewart's network appeared to be more 'optimistic' than Cochrane, and researchers have indicated that they wish for more false positives, than false negatives).

::: callout-note
## Different results

The developed tool may behave differently depending on the Risk of Bias tool that was used. Often, the ROB1 tool was used across all outcomes (rather than individually), and the overall rating for each study is having to be calculated. Therefore, the resulting overall RoBs may be missing important factors that are only specific to certain outcomes (see where chronic rhinosinusitis example is looked at further in section @sec-downgraded-1-level-but-score-of-less-than-1.5). Either the tool only gets used on ROB2 datasets, or it comes with a warning if using ROB1, and maybe the 'adjustment factor' idea can 'fix' these cases.
:::

## Design of RoB element

I've created a function that calculates the weighted average risk of bias, based on the meta-analytic weights used and the overall risk of bias for each study (with respect to the specific outcome where possible). Calculating a weighted average may take into account both of the guidance elements mentioned above (as a high risk study that gives a small weight is unlikely to change the results (depending on how different the estimate is from the pooled result)).

Let's see what the weighted average RoB is for each of our datasets.

```{r, echo = FALSE, output = FALSE}
library(dplyr)
library(purrr)
library(knitr)
library(kableExtra)
source("RoB_downgrades.R")
```

```{r, echo = FALSE}

# Read in info from all datasets
primary_data <- read.csv("datasets.csv")
GRADE_results <- readxl::read_excel("Datasets/primary_info.xlsx", sheet = "GRADE")

# Calculate average for each dataset and add to GRADE dataframe
GRADE_results$RoB_avg <- mapply(function(review_type, version_date, outcome, Rob_tool_used) {
  # subset primary_data
  data <- primary_data %>% 
    filter(Review == review_type & as.Date(Added) <= as.Date(version_date) & Outcome == outcome)
  # run weighted_RoB()
  result <- weighted_RoB(
    data = data,
    rob_tool = Rob_tool_used,
    overall_rob = "Overall.bias",
    random_selection = "random.sequence.generation..selection.bias.",
    allocation_selection = "allocation.concealment..selection.bias.",
    performance = "performance.bias",
    detection = "detection.bias",
    attrition = "attrition.bias",
    reporting = "reporting.bias",
    other = "other.bias",
    weights = "Weight"
  )
  # return average grade
  result
},
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome, GRADE_results$Rob_tool)

# Display results
knitr::kable(GRADE_results %>% select(c("Review", "Version", "Outcome", "Rob_tool", "Downgrade RoB", "RoB_avg")) %>% arrange(RoB_avg), digits = 2)
```

From the above, one can see that at a weighted average RoB score of 1.5, the reviews started getting downgraded by 1 level. None of the example datasets downgraded by two levels for risk of bias, yet some did have the maximum weighted average RoB score of 3.

::: callout-important
## Proposed threshold

Based on the current data, a basic default threshold would be that when the weighted average RoB score is:

-   Greater than or equal to 1.5 - downgrade by one level

-   When to downgrade two levels is less simple and explored further below
:::

In the next sections, I'm going to investigate a bit further the studies that didn't quite follow the general pattern to see if there is anything else quantifiable that can be used to 'fine-tune' the tool.

### Downgraded 1 level, but score of less than 1.5 {#sec-downgraded-1-level-but-score-of-less-than-1.5}

In the chronic rhinosinusitis dataset, for outcomes VAS (disease severity) and avoidance of surgery, the weighted average RoB score was 1 (the lowest possible), yet the authors downgraded the evidence by one level for risk of bias. Note that they used the RoB 1 tool, and assessed it across all outcomes.

For VAS, this was due to "s*tudy limitations: methods or criteria used in the measurement of the outcome were not validated*". In the risk of bias table that is available, the only related information I can find is within the 'selective reporting' field, that there was some lack of clarity regarding the choice of measurement tool, and that for VAS specifically, the outcome was not pre-specified in the original trial registry data. This led to an overall rating of 'Unclear risk' for all 3 studies (note, across all outcomes).

For avoidance of surgery, this was due to "*serious limitations: the criteria used for requiring/not requiring 'rescue surgery' were unclear*". This is similar to the above outcome, in that 'lack of clarity' is mentioned. Specifically, the number of participants who required surgery was missing. As above, this led to an overall rating of 'Unclear' risk for all 3 studies.

As the researchers didn't give an overall risk of bias, my function had to estimate it, and it put all 3 studies as most likely being 'low risk' (note, with 100% probability according to Stewart et. al.'s network).

#### What does this mean?

The big problem here is that the RoB 1 tool doesn't distinguish between different outcomes, and just considers it all together, yet GRADE wants researchers to consider each selection of evidence regarding each specific outcome. This is essentially incompatible.

::: callout-important
## RoB 1 caution

My suggestion is therefore to say that this tool will not perform as well if RoB 1 is being used, due to the above. Researchers can use the proposed tool, but with caution, and will likely have to use the adjustment factor proposed in the below section @sec-adjustment-factor.
:::

### Didn't downgrade, but score of more than 1.5

In the convalescent plasma dataset, for serious adverse events a weighted average RoB score of 1.79 was given for the Feburary 2023 version (which was the same studies in May 2023). In the oral anticoagulation dataset, for outcomes mortality at 6 months and 12 months, a weighted average RoB score of 2.88 and 2.66 was given respectively. Yet for these three outcomes, the researchers did not downgrade the evidence regarding risk of bias.

For serious adverse events in the convalescent plasma dataset, most of the studies were of 'some concerns'. The RoB score for serious adverse events of 1.79 not matching up with downgrading isn't of huge concern and could be categorised as being down to 'difference of opinion' on their significance.

For the mortality outcomes in the oral anticoagulation dataset, the researchers stated that there was "*some concern with risk of bias associated with unclear allocation concealment, taken into account when downgrading from high to moderate certainty*". For both outcomes, the GRADE rating was 'Moderate', downgraded by one level due to serious imprecision. It would seem that the researchers felt this downgrading incorporated the concerns regarding risk of bias, and so no further downgrading was required.

#### What does this mean?

Both scenarios could be corrected with the adjustment factor being proposed below. For convalescent plasma, the 'difference of opinions' between the default settings and the researchers could be adjusted by changing the threshold (going from 1.5 to 1.8 wouldn't be a huge adjustment). For oral coagulation, the researchers are clearly doing their own thing regarding risk of bias, which they can account for by changing the threshold for downgrading (again) due to risk of bias to a much higher value.

### Downgraded only 1 level, but score of 3

There were five instances where all studies have overall high risk of bias, yet the review team only downgraded the evidence by one level due to risk of bias (Colchicine COVID-19 - mortality and worsening; inhaled corticosteroids COVID-19 - symptom resolution; fluvoxamine COVID-19 - serious adverse effects and any adverse effects). The guidelines for downgrading for two levels has the additional factor than any crucial risk identified needs to also likely very seriously alter the results. The standard way to assess this would be to conduct a sensitivity analysis, removing the studies that had high risk of bias. However, in the case where all studies were deemed to be high risk (such as these five examples), this would result in no effect estimate being available, and therefore, no comparison available.

#### What does this mean?

I get the feeling that is is *very* rare for the evidence to be downgraded by two levels due to risk of bias. This is from experience from extracting data (found no examples), and the fact that where there is a high amount of risk of bias, a sensitivity analysis may not be informative (as described above).

::: callout-important
## Default second threshold

Therefore, I think it would be useful to have an option in the tool to state that there is **only one threshold** (i.e., for downgrading one level), and for that option to be the default. If reviewers wish to, they can change this themselves and add a second threshold (that they define) for when to downgrade two levels.
:::

### Adjustment factor {#sec-adjustment-factor}

So far, the idea of an adjustment factor seems sensible as, with the current approach, some research teams downgrade based on risk of bias due to different thresholds/criteria (see sections above); the data collected did not give a straight relationship with the weighted average RoB.

A first approach to this, that is very rudimentary, is to move the default threshold such that the weighted average RoB 'correctly' matches to the set levels the researchers downgraded the outcome for in the most recent version. The new thresholds are then used when considering the review with new studies. This can be thought of as 'calibrating' the tool to the review, and so it'll be focusing on the [likely change]{.underline} from the addition of new studies, rather than a change due to the differences between researchers judgement and tool defaults.

This has been added to the developed function(s) regarding risk of bias.

# Imprecision

The following are suggested ways to quantitatively assess imprecision:

-   Are there enough events?

-   Is the sample size large enough?

-   Does the confidence interval cover multiple messages?

Let's consider each one in a bit more detail regarding practically incorporating it into the tool.

## Are there enough events?

Based on what the smoking review does, they have the following rule based on the total number of events across all arms and all studies:

-   If there are \< 100 events - downgrade 2 levels

-   If there are \< 300 events - downgrade 1 level

However, one does need to consider rare events, and so these thresholds will likely have to be editable. Furthermore, it is only suitable for reviews of dichotomous outcomes. Otherwise, this is a pretty easy method to implement.

## Does the confidence interval cover multiple messages?

This approach is a very practical way of assessing imprecision by actively seeing if the review is suggesting multiple 'messages' (i.e., clinical harm and benefit), a direct consequence of imprecision. It does so by the following categories:

-   If the confidence interval (CI) includes a meaningfully positive or negative effect, as well as no meaningful effect in the opposite direction, then downgrade 1 level

-   If the CI includes both meaningfully positive and negative effects, then downgrade 2 levels

The main element that needs defining in this approach is what value translates to a 'meaningful' effect? The following defaults have been suggested by the literature:

-   For dichotomous outcomes, a meaningful effect is when the relative risk reduction or increase is greater than 25%

-   For continuous outcomes, the minimal importance difference should be used if known, otherwise a default of +/- 0.5 can be used.

Again, as with considering number of events, these thresholds need to be editable by the reviewers where appropriate. This includes where the review topic is considering rare events.

## Is the sample size large enough?

This has been suggested by multiple sources. For dichotomous outcomes, it has been suggested to consider whether the total sample size across all trials included in the review is lower than an optimal information size. If so, downgrade 1 level. However, how should the optimal information size be calculated? It is often suggested to calculate the value as if it is a single trial and compare, but there are equations out there for calculating the optimal number of trials and average trial sample size for a meta-analysis. For continuous outcomes, a 'rule of thumb' that has been suggested is that is there are \< 400 participants in total across the review, then downgrade 1 level.

Due to the discrepancies around how to calculate an optimal information size, I'm tempted to not utilise this approach. Additionally, I believe the other two approaches above probably cover any imprecision that would be 'found' by considering the total sample size of the review.

## Testing tool

We have two approaches for downgrading, both of which can lead to downgrading 2 levels. However, the tool can only downgrade a maximum of 2 levels (as per GRADE guidance). Where both approaches are appropriate (i.e., for dichotomous outcomes), the maximum number of levels downgraded will be taken. 'Summing' them doesn't make sense as both approaches are attempting to 'measure' the same thing, and so they shouldn't accumulate. Taking the maximum was preferred over taking the minimum as it ensures that any information indicative of imprecision isn't ignored. This approach also follows how the e-cigarettes review team assess imprecision.

Let's see how well the tool performs at predicting the imprecision grading levels using the default values suggested by the literature.

```{r, echo = FALSE, output = FALSE}
source("Imprecision_downgrades.R")
```

```{r, echo = FALSE}
# Calculate total events and extract confidence intervals for each dataset and add to GRADE dataframe
GRADE_results$total_events <- mapply(function(review_type, version_date, outcome, measure) {
  if (measure == "RR") {
    # subset primary_data
    data <- primary_data %>% 
      filter(Review == review_type & as.Date(Added) <= as.Date(version_date) & Outcome == outcome)
    # run sum_events()
    sum_events(
      data = data,
      event_cols = c("Experimental.cases", "Control.cases")
    )
  } else {
    NA
  }
},
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome, GRADE_results$Measure)

GRADE_results$CI <- mapply(function(review_type, version_date, outcome, measure, model) {
  # subset primary_data
  data <- primary_data %>% 
    filter(Review == review_type & as.Date(Added) <= as.Date(version_date) & Outcome == outcome)
  # run meta-analysis
  meta <- metafor::rma(data = data, yi = yi, vi = vi, measure = measure, method = model)
  if (measure == 'RR') {
    paste0(round(exp(meta$ci.lb), 2), ", ", round(exp(meta$ci.ub), 2))
  } else {
    paste0(round(meta$ci.lb, 2), ", ", round(meta$ci.ub, 2))
  }
}
,
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome, GRADE_results$Measure, GRADE_results$Model)

# Predict downgrading levels based on the information
GRADE_results$imprecision_levels <- mapply(function(total_events, CI, measure) {
  
  # predict number of levels to downgrade by
  imprecision_downgrades(
    outcome = measure, total_events = total_events, CI_lb = log(as.numeric(stringr::str_split_fixed(CI, ",", n = 2)[1])), CI_ub = log(as.numeric(stringr::str_split_fixed(CI, ",", n = 2)[2]))
  )

},
GRADE_results$total_events, GRADE_results$CI, GRADE_results$Measure)

# Display results
knitr::kable(GRADE_results %>% select(c("Review", "Version", "Outcome", "Downgrade Imprecision", "total_events", "CI", "imprecision_levels")) %>% arrange(total_events), digits = 2)
```

## Exploring differences

Of the 47 outcomes, the tools prediction of the number of levels to downgrade matched the reviewers assessment for `{r} sum(GRADE_results[, "Downgrade Imprecision"] == GRADE_results$imprecision_levels)` outcomes. Let's explore those that didn't match and see if we can know the reason.

### Instances where tool over-downgraded

-   Chronic rhinosinusitis - serious adverse events & avoidance of surgery

    -   SAEs: 67 events; 0.3 to 0.8. Avoidance: 27 events; 0.05 to 0.52. The reviewers only rated down 1 level for serious adverse events and none for avoidance, whereas the tool suggested downgrading two levels for both outcomes due to both having less than 100 events.\
        For SAEs, the reviewers recognised that there was a small sample size, but didn't think it was worth downgrading two levels - however, it doesn't make sense that the lower sample size for avoidance didn't have any downgrading due to imprecision (the outcome was only downgraded due to risk of bias).

-   Convalescent plasma - adverse events & serious adverse events

    -   AEs: 69 events; 0.58 to 1.41. SAEs: 577 events; 0.91 to 1.43. The reviewers only rated down 1 level for AEs and none for AEs, whereas the tools suggested two levels for AEs due to the very small number of events and wide CI, and one level for SAEs due to the wide CI. The review authors acknowledge the few participants and wide intervals for AEs, but still only offered to downgrade 1 level.

-   Ivermectin - adverse events & viral clearance

    -   AEs: 47 events; 0.61 to 1.79. Viral clearance: 97 events; 0.8 to 1.58. The reviewers only rated both outcomes down one level, whereas the tool suggested 2 levels due to less than 100 events in both outcomes (and wide CI for AEs). The review authors acknowledges the few participants and wide CIs, but still didn't downgrade as far as the tool did.

-   Systematic corticosteroids - mortality at 120 days

    -   186 events; 0.24 to 2.3. The reviewers only rated down one level, whereas the tool suggested 2 levels due to the hugely wide CI. The reviewers acknowledged the wide CI, but didn't downgrade to the degree that the tool did.

-   Oral anticoagulation - major & minor bleeding

    -   Major: 143 events; 2.05 to 4.09. Minor: 143 events; 1.89 to 5.02. The review authors didn't downgrade for imprecision, but the tool did by one level due to having less than 300 events. Naturally, the authors won't give a reason for *not* downgrading, but I imagine it's likely to do with the CIs being very much in the positive 'zone' (i.e., far away from region of 'no effect'). **Does this counteract the small number of events?** The authors could change the threshold if 143 isn't concerning for them - but I don't think one 'measure' of imprecision should interact with the other.

-   Remdesivir - serious adverse events

    -   634 events; 0.65 to 1.07. The review authors didn't downgrade for imprecision, but the tool did for having a wide CI. Did the review authors not think the position of the CI was a concern because of the large number of events?

-   Inhaled corticosteroids - symptom resolution

    -   1,009 events; 1.09 to 1.3. Same as remdesivir above. Furthermore, the CI was only just beyond the default threshold for clinical significance.

-   Janus kinase - mortality at 28 and 60 days & serious adverse events

    -   Mortality28: 1,346 events; 0.57 to 0.91. Mortality60: 249 events; 0.56 to 0.86. SAEs: 530 events; 0.68 to 0.92. The review authors didn't downgrade any of the outcomes for imprecision, but the tool downgraded each for 1 level: due to wide CI for all, and also due to number of events for mortality at 60 days. Apart from mortality at 60 days, the other outcomes had very high numbers of events. Furthermore, whilst the CI were deemed wide due to the default values, there were all one side of the null effect. As such, the review authors may have had a different opinion on clinical significance.

-   Antibiotics - serious adverse events & arrhythmias

    -   SAEs: 185 events; 0.89 to 1.4. Arrhythmias: 341 events; 0.73 to 1.15. The reviewers didn't downgrade for precision, whereas the tool did by one level due to a wide CI for both events (and low number of events for SAEs). The wide CIs were borderline regarding the default value.

From assessing these case-studies, there appears to be two general themes appearing:

1.  Review authors have different thresholds (in both number of events, and wide CIs) of what they consider is worth downgrading (which is erring on the side of making the data look 'better'). Regarding the wide CIs, I'm not entirely sure whether some of them have considered a null effect *region* rather than just a line.

2.  If one of the imprecision measures seem to indicate no problems, reviewers may feel that it invalidates potential evidence of imprecision from the other measure. E.g., when the number of events is very high, but the CI is still wide enough to cover two clinically significant states. This shouldn't be the case - lack of evidence doesn't **prove** that things are okay.

### Instances where the tool under-graded

-   Fluvoxamine - adverse events

    -   201 events; 0.82 to 1.37. The reviewers downgraded by two levels, whereas the tool only downgraded by one level (due to \< 300 events and CI going beyond 1.25, but not beyond 0.75). The reviewers downgraded by 2 levels "since the CI suggests both potential benefit and potential harm". Unfortunately I couldn't find any details regarding what threshold they considered to be clinical benefit/harm - disagreement seems related to this.

-   Oral anticoagulation - mortality at 12 months

    -   717 events; 0.87 to 1.03. The tool didn't downgrade for imprecision but the reviewers did by one level, with the reason being "confidence interval of absolute effect suggests both potential harm and potential benefit, with large number of events". No other details available regarding threshold used.

-   Remdesivir - mortality & adverse events (Jan '23 update only)

    -   Mortality: 749 events; 0.81 to 1.06. AEs: 1,463 events; 0.92 to 1.18. The tool didn't downgrade for imprecision but the reviewers did for both outcomes by one level, with the reason being "wide confidence intervals in the studies and/or the 95% confidence interval includes both benefits and harms". Again, no details of threshold used.

In summary, the only reason these cases didn't agree with the tool is that they were using a different threshold for defining clinical benefit and harm. Unfortunately, the thresholds that they used were not reported (in the main text at least). However, this doesn't impact the functionality of the tool as users can specify their own thresholds - which these LSRs indicate they have.

::: callout-caution
## Adjustment factor now optional

Whether the tools should adjust the thresholds based on previous reviews is now an option, rather than doing so by default. In an ideal world, reviewers should be competent enough to have a suitable system for carrying out GRADE, and so should be able to input their own thresholds that should allow the tool to marry up with previous versions.

I suggest only minor cases for using the adjuster. These cases would be for myself to test it on reviews for which I don't have contact with the authors, and potentially reviews that are midway through being living, and wish to have some help with defining the thresholds. Therefore, the tool will also output which thresholds it used (so that users can be confident in knowing what was used).
:::

::: callout-note
## Observation

Nicola Lindson suggested some reviewers might not downgrade as far as the proposed tool due to fears of it looking badly on their review. This might be more prevalent when the outcome has already been downgraded for a different domain (i.e., the reviewers might be hesitant to further downgrade it regarding another domain).

Can I quantify this?
:::

For instances when the adjustment factor is needed, here it is less simple to implement as we have two approaches for dichotomous outcomes. I have created an auto-adjust matrix that decides on how and which thresholds are altered depending on how there is disagreement (i.e., is it by one measure, or both, and in which directions). Generally, from experience from comparing disagreements, when both measures disagreed, the reviewers specified it was regarding the CI. Therefore, when both measures had disagreement but only one measure was needed to gain overall agreement (as the maximum is taken), often the threshold relating to CI was adjusted. I created a matrix of which thresholds to adjust and in which direction depending on where the disagreements lie.

# Indirectness

There are two elements of indirectness that can occur: a difference between studies included and review aims, and when indirect comparisons are utilised due to lack of direct evidence[@guyatt2025]. As we're only considering pairwise meta-analyses and RCTs, for now we're going to assume that the reviewer has not needed to find indirect evidence, and so we will focus on the former element of indirectness. Otherwise, the GRADE guidance for authors suggests considering whether the evidence found is more restrictive than the review question in terms of PICO[@schünemann2013]. Due to the nature of this domain, there will be no quantitative elements, and instead, the tool will have the following prompt questions from the user to answer.

-   Are the populations presented in the current evidence so much more restrictive than the target population of the review that they're likely to affect the magnitude of effect substantially? If so, is it serious or very serious?

-   Are the interventions that have been studied in the current pool of evidence more restrictive than the target interventions of the review? If so, is it serious or very serious?

-   Are the comparators that have been included in the current studies more restrictive than the target comparator(s) of the review? If so, is it serious or very serious?

-   Are the outcomes that are presented in the current evidence more restrictive than the target outcomes of the review? If so, is it serious or very serious?

The questions are not to be answered on their own but as a collective as the judgement of whether to downgrade due to indirectness is to be done based on the combined effect of the areas considered above[@guyatt2011]. As such, the user will simply be asked to judge the evidence based on the prompt questions, as to whether any problems identified need to be rates down by two, one, or no levels. Additionally, when presented with the proposed tool, reviewers will also be given links to guidelines written by GRADE developers[@guyatt2011; @guyatt2025a]

Unfortunately, due to the nature of this domain, I cannot 'test' it, or measure it against any of the reviews I've collated.

As the judgement of indirectness requires knowledge of the study characteristics, this judgement should be able to be made in-between data extraction and committing to an update. From experience, some of the research team will be aware of the characteristics of the new studies, and whether their accumulating changes any levels of indirectness.

# Inconsistency

Both Stewart et. al.[@stewart2015], the GRADE guidance[@schünemann2013a], and Cochrane author advice videos encourage researchers to consider the following elements when making a judgement regarding inconsistency:

-   Is there a wide variation in effect estimates?

-   Do the confidence intervals overlap?

-   Statistical test for heterogeneity (threshold for p-value needed)

-   $I^2$ statistic, with the following rough guide:

    -   0% to 40% - low heterogeneity

    -   30% to 60% - moderate heterogeneity

    -   50% to 90% - substantial heterogeneity

    -   75% to 100% - considerable heterogeneity

It should be notes that before considering any of the above, authors should have already considered whether any heterogeneity can be adequately explained (and adjusted for). As such, the reviewer should be making a judgment based on any remaining heterogeneity/inconsistency.

Whilst $I^2$ is a simple quantitative measure, due to ambiguous interpretation and researchers in the field suggested is may be misleading[@guyatt2025b], I don't think this would be a good element to use in the tool. For similar reasons, I don't think utilising a test statistic for heterogeneity would be the best approach. Core GRADE[@guyatt2025b] considers a visual inspection of the forest plot to assess is there is any overlap of confidence intervals, difference in point estimates (with a focus on a chosen threshold representing the null effect or minimal important difference). As such, the tool will try to incorporate the first two elements in the list above.

We will follow the plan that Core GRADE suggests[@guyatt2025b], which considers the following three questions:

1.  Are there important differences in point estimates of studies and limited overlap of their CIs? If no, then don't rate down and stop. If yes, consider next question.

2.  Evaluate point estimates of studies in relation to chosen threshold. If the majority are on one side, then don't rate down and stop. If there are substantial proportions on both sides, then consider the next question.

3.  Is the current inconsistency explained by credible subgroup analyses? If not, then rate down for inconsistency. Otherwise, present results separately for each subgroup and evaluate inconsistency separately.

The tool will roughly follow the same plan (i.e., it should give the same results), but have a slightly different order of operations.

## Overlap of CIs

We will first consider how much the confidence intervals are overlapping. The reason for not starting with 'important' differences in point estimates, is that this is hard to categorise. Furthermore, overlapping of confidence intervals and variation regarding a user-defined threshold will likely cover presence of inconsistency.

I asked ChatGPT (GPT-5) for help coming up with ideas based on mathematical principles for quantitatively assessing the overlap of 95% CIs. The general idea it came up with was to calculate the 'common probabilistic overlap' where each CI is represented by the normal distribution is was estimated from. I asked it for code and references, but this did give some packages that don't exist and non-existing references - but it gave me an idea to start from (and some legit references to look at).

::: callout-note
## ChatGPT prompts

1.  Does there exact a mathematical way to quantitatively assess overlap of multiple ranges?

2.  Are there any mathematical extensions to the problem, given that the values within each range aren't equal (i.e., the ranges are actual confidence intervals, and so I'd be more interested in overlap of values nearer the middle, than the edges)

3.  I'm looking at the overlap of 95% confidence intervals, which dont have a pdf, but use a pdf to calculate the interval. Are there any applications to 95% confidence intervals specifically?

4.  Can you show me how to compute the probabilistic overlap of three 95% confidence intervals using R code please

5.  Do you have a reference for calculating the common probabilistic overlap?

6.  Can you extend the Jaccard Index two multiple ranges, and give an example in R code (this was suggested after the first prompt)

7.  Do you have a reference for the mean pairwise Jaccard index?
:::

Using probability distributions proved to be too difficult - so let's start simple and assume uniform distribution (and so is a geometric overlap, rather than a statistical one).

The Jaccard index is an option, but the standard extension to multiple ranges, defining a strict intersection, is not suitable as the overlap may not be one simple range, and there may be no place where **all** confidence intervals overlap, but a few might overlap with each other. ChatGPT suggests an average pairwise Jaccard index as this gives a better sense of overall similarity.

The Jaccard index is the intersection of sets divided by the union of sets[@real1996]:

$$
J(S_a, S_b) = \frac{|S_a \cap S_b|}{|S_a \cup S_b|}
$$

When there are more than two sets, when interested in the overall similarity, rather than the strict overlap, the pairwise average Jaccard Index can be calculated[@fan2016]:

$$
\bar{J} = \frac{2}{n(n-1)} \sum_a \sum_{b \ne a} J(S_a, S_b)
$$

Calculate the intersection and union of two sets of ranges:

$$
CI_a = [LB_a, UB_a], CI_b = [LB_b, UB_b]
$$

$$
|CI_a \cap CI_b| = max(0, min(UB_a, UB_b) - max(LB_a, LB_b))
$$

$$
|CI_a \cup CI_b| = max(UB_a, UB_b) - min(LB_a, LB_b)
$$

However, confidence intervals can be negative (especially when on the log scale, which would be optimal for ratios, otherwise they wouldn't be symmetrical). To take into account the fact that values can be negative, we will transform all the CIs to be positive (this will still give an equivalent Jaccard Index as its standardised). To do this, we'll take the smallest lower bound, and if it's negative, add the modulus of that value to all upper bounds and lower bounds. Otherwise, when the intersection of the two intervals lies in the negative space, there is no way to differentiate the legit intersection from no intersection.

## Variation in point estimates regarding threshold

Based on a designated value of importance, simply calculate the proportion of study estimates that were on one side of the threshold (and try and find a default based on the data we have here). The default value will simply be the null value (0 for differences, 1 for ratios).

Lets calculate the average pairwise Jaccard index and variation proportion for each review outcome.

```{r, echo = FALSE, output = FALSE}
source("Inconsistency_downgrades.R")
```

```{r, echo = FALSE}
# Calculate pairwise Jaccard index for each dataset and add to GRADE dataframe
GRADE_results$Jaccard <- mapply(function(review_type, version_date, outcome) {
  # subset primary_data
  data <- primary_data %>% 
    filter(Review == review_type & as.Date(Added) <= as.Date(version_date) & Outcome == outcome & Weight != 0)
  # run pairwise_Jaccard()
  pairwise_Jaccard(
    CI_lb_col = as.numeric(data$`CI.start`),
    CI_ub_col = as.numeric(data$`CI.end`)
  )
},
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome)

# Calculate proportion of estimates that are either side of 0
GRADE_results$est_var <- mapply(function(review_type, version_date, outcome, measure) {
  # subset primary_data
  data <- primary_data %>% 
    filter(Review == review_type & as.Date(Added) <= as.Date(version_date) & Outcome == outcome & Weight != 0)
  # run estimate_variation() 
  estimate_variation(
    estimates = data$yi
  )
},
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome, GRADE_results$Measure)

# Display results
knitr::kable(GRADE_results %>% select(c("Review", "Version", "Outcome", "Downgrade Inconsistency", "Jaccard", "est_var")) %>% arrange(est_var, Jaccard), digits = 2)
```

Generally lower Jaccard values to match to reviews that downgrading for inconsistency, however, they are not 'clumped' together to give a satisfactory default threshold. Looking at how the estimates varied didn't give a clear answer either, all studies with the lowest score (50%) didn't downgrade, yet one study with the highest score (100%) did downgrade. There could be reviewer differences happening, in particular regarding number of studies and other thresholds.

### Exploring instances where downgraded for inconsistency

To give some more context to the above results, let's take a closer look at the outcomes that were downgraded by 1 level for inconsistency by the reviewers.

-   Systemic corticosteroids for COVID-19 on mortality as 120 days. 67% of studies were on one side of the threshold, and a Jaccard index of 0.03. Review just simply says they downgrading it due to inconsistency, with no details on specifics.

-   Antibiotics for COVID-19 on improvement. 67% of studies were on one side of the threshold, and a Jaccard index of 0.18. Reason was specified: "... three studies contributed heterogeneous results (I squared = 49%). One study favoured standard of care alone and the other two showed no difference between study arms."

-   Remdesivir for COVID-19 on adverse events (August 2021 version). 67% of studies were on one side of the threshold, and a Jaccard index of 0.24. Stated that downgraded "because of statistical heterogeneity (I squared = 77%)".

-   Adverse events regarding treatment for non-small cell lung cancer. All studies were on one side of the threshold, but had a Jaccard index of 0.34. Only detail found was "...because of inconsistency of results and a high heterogeneity score".

From looking at the Cochrane reviews obtained for this exercise, it doesn't seem that they can be used to set default values. It seems that I squared values were most often use to assess inconsistency, however, using this solely is outdated and always comes with warning - the core GRADE methods focus on overlap and variation instead. Furthermore, it seems to be subjective as there were no patterns (as partially seen with other domains) and some cases where they weren't downgraded really seem like they should have been based on the scores (e.g., 50% of studies on one side and a Jaccard index of 0.07).

Based on subjective instinct, I'll suggest the below defaults.

::: callout-important
## Proposed thresholds

From experience, reviews only ever downgraded by one level. As the variation in estimates is more indicative of a change in conclusions, I propose the following threshold structure:

-   If the variation in estimates is 0.8 or higher (i.e., fewer than 1 in 5 'disagree'), then don't downgrade for inconsistency (i.e., the level of overlap doesn't matter). If it is below 0.8, then consider the overlap:

    -   If the Jaccard index is lower than 0.4, then downgrade by one level for inconsistency (based on previous reviews).

The autoadjust option will work by first adjusting the estimate variation, and then the Jaccard index.
:::

## Subgroup consideration

GRADE guidance suggests that when inconsistency is due to known variables, that the results should be presented by such subgroupings, and inconsistency to be assessed within each group. There isn't a subgroup option within the inconsistency function itself, however, when combining all the functions together to create the 'main' function, a subgroup option will be available (as it's assumed that the reviewers already know of variables that result in inconsistency).

# Publication bias

The following criteria were encouraged by both Stewart et al[@stewart2015] and GRADE guidance[@schünemann2013]:

1.  Discrepancies between published and unpublished data

2.  Industry influence

3.  Search integrity

4.  Funnel plot statistics

The first three points are unlikely to change across a living review and are very nuanced, as such, they will simply be yes/no questions as per indirectness. To make a start, we'll just look at funnel plot statistic where there are 5 or more studies. We can use either the Cochrane examples of Stewart et al[@stewart2015] network thresholds to define a default threshold.

Let's see what the Egger's test statistic is for each dataset we have.

```{r, echo = FALSE}
# Calculate Egger's test for each dataset and add to GRADE dataframe
GRADE_results$Eggers <- mapply(function(review_type, version_date, outcome) {
  # subset primary_data
  data <- primary_data %>% 
    filter(Review == review_type & as.Date(Added) <= as.Date(version_date) & Outcome == outcome & Weight != 0)
  # Check there are at least 5 studies
  n <- nrow(data)
  if (n < 5) {
    NA
  } else {
    # Calculate funnel plot statistic (Egger's)
    res <- metafor::rma(data = data, yi = yi, vi = vi)
    test_info <- metafor::regtest(res, model = "lm", predictor = "sei")
    test_info$pval
  }
},
GRADE_results$Review, GRADE_results$Version, GRADE_results$Outcome)

# Display results
knitr::kable(GRADE_results %>% select(c("Review", "Version", "Outcome", "Downgrade Publication bias", "Eggers")) %>% arrange(Eggers), digits = 2)
```

Of the reviews that did have 5 or more studies, all of those with a p-value below 0.85 weren't downgraded, but one that was above 0.85 (0.86) was downgraded for publication bias. This was the topic of serious adverse events for convalescent plasma (Feb 2023 version). Normally you would expect a threshold of 0.95, but the reviewers stated they suspected publication bias "...because safety outcomes were assessed and reported in most studies for convalescent plasma group only" - so this was due to one of the other factors (they state that they intended to use p \< 0.1 as significance). Adverse events was also downgraded for publication bias even though it had less than 5 studies (May 2021 version) due to the same reasons as above.

We will set the **default threshold to be a p-value of 0.9**. However, the final function needs to be able to incorporate downgrading for the other trigger questions.

# References

::: {#refs}
:::
