---
title: "Interpretation Tool"
format: docx
editor: visual
---

This document is to go over the details of the interpretation tool part of the update framework, which is based on the work by Langan et. al.

```{r setup, include = FALSE}

# Load function and example datasets
source("InterpretationTool.R")

# Remove progress bars when rendering
if (!interactive()) {
  progressr::handlers("void")
}

### Test Data antibiotics vs. control for the common cold to alleviate symptoms by 7 days ###
raw_data_bin <- data.frame(StudyID = c(1, 2, 3, 4, 5, 6), Study = c("Herne_1980", "Hoaglund_1950", "Kaiser_1996", "Lexomboon_1971", "McKerrow_1961", "Taylor_1977"),
                   R.1 = c(7, 39, 97, 8, 5, 12), N.1 = c(7+39, 39+115, 97+49, 8+166, 5+10, 12+117), T.1 = rep("Treatment", 6),
                   R.2 = c(10, 51, 94, 4, 8, 3), N.2 = c(10+12, 51+104, 94+48, 4+83, 8+10, 3+56), T.2 = rep("Control", 6))
raw_data_new <- data.frame(StudyID = c(7), Study = c("Fake_1"),
                           R.1 = c(63), N.1 = c(63+252), T.1 = rep("Treatment", 1),
                           R.2 = c(77), N.2 = c(77+105), T.2 = rep("Control", 1))
raw_data_new_multiple <- data.frame(StudyID = c(7, 8, 9), Study = c("Fake_1", "Fake_2", "Fake_3"),
                           R.1 = c(63, 12, 45), N.1 = c(63+252, 12+34, 45+60), T.1 = rep("Treatment", 3),
                           R.2 = c(77, 21, 58), N.2 = c(77+105, 21+25, 58+42), T.2 = rep("Control", 1))

### Obtain study effects and standard errors #
MAdata_bin <- metafor::escalc(measure = 'RR', ai = R.1, bi = N.1-R.1, ci = R.2, di = N.2-R.2, data = raw_data_bin)   # gives ES (effect estimate) and seES (sampling variances) on logOR scale for binary data
MAdata_bin_new <- metafor::escalc(measure = 'RR', ai = R.1, bi = N.1-R.1, ci = R.2, di = N.2-R.2, data = raw_data_new)
MAdata_bin_new_multiple <- metafor::escalc(measure = 'RR', ai = R.1, bi = N.1-R.1, ci = R.2, di = N.2-R.2, data = raw_data_new_multiple)
MAdata_bin$sei <- sqrt(MAdata_bin$vi)  # Calculate standard errors
MAdata_bin_new$sei <- sqrt(MAdata_bin_new$vi)
MAdata_bin_new_multiple$sei <- sqrt(MAdata_bin_new_multiple$vi)
```

# Derivation of contours

This section will go over the derivation of forming the contours of the plot for different situations.

## Fixed-effects

We are working within the fixed-effects framework initially, where the between-study heterogeneity, denoted by $\tau^2$, is equal to zero.

### Statistical significance

Lets assume we are interested in changing the statistical significance (i.e., the p-value).

#### One new study

Let's start with assuming that there is only one new study that will be added to the meta-analysis.

The 95% confidence interval (CI) is denoted by

$$\hat{\theta} \pm \Phi^{-1} (\frac{\alpha}{2}) \sqrt{var(\hat\theta)}$$ {#eq-ci}

where $\hat\theta$ is the pooled effect estimate, $\alpha$ is the type I error rate (normally set at 0.05), $\Phi^{-1}$ is the inverse cumulative normal distribution with mean 0 and variance 1 (which ensures that $\Phi^{-1}(\alpha/2)$ maps an $\alpha$ value of 0.05 to the value of 1.96). $var(\hat\theta)$ is the variance of the pooled effect estimate.

To create a threshold for statistical significance (specified by $\alpha$), we can specify that the lower or upper bound of the 95% CI needs to touch the null value, denoted by $\theta_0$.

Let's assume that there is a new study with (unknown) effect estimate $\hat\theta_{new}$ , variance $var(\hat\theta_{new})$, and weight $w_{new}$. Then once this new study is added to the meta-analysis, the updated results will be as follows:

$$
\hat\theta = \frac{\sum^{n}_{i=1}{w_i\hat\theta_i} + w_{new}\hat\theta_{new}}{\sum^n_{i=1}{w_i} + w_{new}}
$$ {#eq-updated-estimate}

$$var(\hat\theta) = \frac{1}{\sum^n_{i=1}{w_i} + w_{new}}$$ {#eq-updated-variance}

where $n$ is the number of studies in the original meta-analysis.

By substituting @eq-updated-estimate and @eq-updated-variance into @eq-ci, and specifying that the upper or lower bound equals $\theta_0$, we can get an expression for $\theta_{new}$ in terms of other quantities:

$$
\hat\theta_{new} = \frac{1}{w_{new}}\{\theta_0[\sum^n_{i=1}{w_i} + w_{new}] - \sum^n_{i=1}{w_i \hat\theta_i} \pm \Phi^{-1}(\frac{\alpha}{2}) \sqrt{\sum^n_{i=1}{w_i} + w_{new}}\}
$$ {#eq-rearranged-theta-new}

In a fixed-effects meta-analysis, each weight is simply calculated as

$$
w_i = \frac{1}{var(\hat\theta_i)}
$$ {#eq-fixed-weight}

Therefore, by substituting @eq-fixed-weight into @eq-rearranged-theta-new, we then get an equation for the contour threshold on the estimate-variance plane.

##### Example plot

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'fixed', outcome = 'RR', zero = 1)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

#### Two new studies

When expanding from one new study to two new studies, @eq-updated-estimate and @eq-updated-variance become

$$
\hat\theta = \frac{\sum^{n}_{i=1}{w_i\hat\theta_i} + w_{n1}\hat\theta_{n1} + w_{n2}\hat\theta_{n2}}{\sum^n_{i=1}{w_i} + w_{n1} + w_{n2}}
$$ {#eq-meta-theta-2studies}

$$var(\hat\theta) = \frac{1}{\sum^n_{i=1}{w_i} + w_{n1} + w_{n2}}$$ {#eq-meta-variance-2studies}

where $\hat\theta_{nj}$ and $w_{nj}$ are the estimates and weights of new studies respectively, where $j = 1,2$.

Let

$$w_{\bar{new}} = w_{n1} + w_{n2}$$ {#eq-weight-2studies}

$$\hat\theta_{\bar{new}} = \frac{w_{n1}\hat\theta_{n1} + w_{n2}\hat\theta_{n2}}{w_{n1} + w_{n2}}$$ {#eq-theta-2studies} Substituting @eq-weight-2studies and @eq-theta-2studies into @eq-meta-theta-2studies and @eq-meta-variance-2studies gives us the following:

$$
\hat\theta = \frac{\sum^{n}_{i=1}{w_i\hat\theta_i} + w_{\bar{new}}\hat\theta_{\bar{new}}}{\sum^n_{i=1}{w_i} + w_{\bar{new}}}
$$

$$
var(\hat\theta) = \frac{1}{\sum^n_{i=1}{w_i} + w_{\bar{new}}}
$$

This is simply the same equations as equations @eq-updated-estimate and @eq-updated-variance. Therefore, everything else follows through as before, but with $w_{\bar{new}}$ and $\hat\theta_{\bar{new}}$ in place of $w_{new}$ and $\hat\theta_{new}$.

#### Multiple new studies {#sec-multiple-new-studies-fixed}

Where we are calculating the contour to $m$ new studies, the equations follow as if it was one new study, but $w_{new}$ and $\hat\theta_{new}$ are defined as

$$
w_{\bar{new}} = \sum^m_{j=1}{w_{nj}}
$$

$$
\hat\theta_{\bar{new}} = \frac{\sum^m_{j=1}{w_{nj}\hat\theta_{nj}}}{\sum^m_{j=1}{w_{nj}}}
$$

where $\hat\theta_{nj}$ and $w_{nj}$ are the estimates and weights of the new studies respectively, for $j = 1,2,...,m$.

##### Example plot

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'fixed', outcome = 'RR', zero = 1, sig_level = 0.0005)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

### Specific confidence interval bounds

Let's suppose that instead of wanting a statistically significant pooled estimate, that the researchers are looking for the 95% CI bounds to be at a certain value (likely of clinical importance).

With statistical significance, the contour equations are calculated from stating that the upper or lower bounds must touch the null value, $\theta_0$. In this scenario, the contour equations would be calculated from the following thresholds:

$$
\theta_{lower} = \hat{\theta} - \Phi^{-1} (\frac{\alpha}{2}) \sqrt{var(\hat\theta)}
$$ {#eq-lower-bound}

$$
\theta_{upper} = \hat{\theta} + \Phi^{-1} (\frac{\alpha}{2}) \sqrt{var(\hat\theta)}
$$ {#eq-upper-bound}

where $\theta_{lower}$ and $\theta_{upper}$ are the user-defined lower and upper 95% CI bounds of interest.

Substituting @eq-updated-estimate and @eq-updated-variance into @eq-lower-bound and @eq-upper-bound, and rearranging gives the following expression(s) for $\hat\theta_{new}$.

$$
\hat\theta_{new} = \frac{1}{w_{new}}\{\theta_{lower}[\sum^n_{i=1}{w_i} + w_{new}] - \sum^n_{i=1}{w_i \hat\theta_i} + \Phi^{-1}(\frac{\alpha}{2}) \sqrt{\sum^n_{i=1}{w_i} + w_{new}}\}
$$ $$
\hat\theta_{new} = \frac{1}{w_{new}}\{\theta_{upper}[\sum^n_{i=1}{w_i} + w_{new}] - \sum^n_{i=1}{w_i \hat\theta_i} - \Phi^{-1}(\frac{\alpha}{2}) \sqrt{\sum^n_{i=1}{w_i} + w_{new}}\}
$$

Whilst in the fixed-effects framework, the extensions to multiple new studies is the same here as for statistical significance.

##### Example plot

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'fixed', outcome = 'RR', upper = 0.9)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

## Random-effects

When working in the random-effects framework things get more complex as $\tau^2 \ne 0$ and is estimated from the studies in the meta-analysis, including the new studies.

Let's start again with only considering one new study and aiming for a meta-analysis with statistical significance at the $\alpha = 0.05$ level.

Under a random-effects framework, the weight for each study (including new studies) is instead calculated as

$$
w_i = \frac{1}{var(\hat\theta_i) + \tau^2_{new}}
$$ {#eq-random-weight}

where $\tau^2_{new}$ is the estimated between-study heterogeneity when included the new studies into the meta-analysis. Unfortunately, we can't simply plug the random-effects equation for weights into the contour equations derived for fixed effects (@eq-rearranged-theta-new) as $\tau^2_{new}$ depends on $\hat\theta_{new}$. Furthermore, we can't simply plug in an equation for $\tau^2_{new}$ and rearrange as calculating an estimate for between-study heterogeneity is too complicated to do by hand, and so researchers often use established estimators instead, such as the DerSimonian-Laird estimator or Restricted Maximum Likelihood. (Note the further challenge, that these commonly used estimators wouldn't allow us to incorporate the uncertainty in estimating between-study heterogeneity).

As such, an equation to form the contour thresholds cannot be derived as was done for the fixed-effects framework. Instead, as per Langan et. al., the standard error-estimate plane is broken down into 'pixels', where $\tau^2_{new}$ is estimated at each pixel (calculated from the original studies as well as the new study represented by the pixel). For each pixel, $\tau^2_{new}$, $\hat\theta_{new}$, and $var(\hat\theta_{new})$ are plugged into the meta-analysis equations ( @eq-updated-estimate and @eq-updated-variance), but using the weights calculated in @eq-random-weight, and noted whether or not the threshold was crossed (i.e. whether 95% CI was one side of null effect). The result at each pixel is coloured according to whether the threshold was met. This easily extends to the clinical bound thresholds as well as the statistical significance.

##### Example plot

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 100)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

### Multiple new studies {#multiple-new-studies-1}

Under the random-effects model, we have to take the estimate-variance plane 'pixel-by-pixel'. When there was one new study, this was simple as the pixel had one set of possible values 'attached' to it that were easily obtained. When considering multiple new studies, the idea of using the 'average' (using random-effects weightings) of the new studies (as was done for fixed-effects) doesn't work for multiple reasons. Firstly, for each pixel, it is not possible to know what new studies created the 'average' represented by the respective point. Without knowing the new studies, the between-study heterogeneity can't be estimated. Secondly, different combinations of new studies could lead to the same 'average' effect size and standard error. It therefore follows that as there is no unique combination, there could be multiple different estimates of the between-study heterogeneity represented by a single pixel. To move forward, a compromise or assumption has to be made.

One idea is to assume a constant between-study heterogeneity across all pixels in the plane. If this idea was to be carried out, then the optimal choice for the common between-study heterogeneity across all pixels would be the updated meta-analysis, $\tau^2_{new}$. Whilst the new studies 'average' would at least sit in the correct section of the plot, this becomes a very strong assumption as you get to more 'extreme' pixels/values. This has a consequence on the contours created, as the further extreme you go in the plot, the between-study heterogeneity is likely being underestimated, and therefore being coloured 'incorrectly' as leading to significance, whereas such cases would be unlikely.

::: callout-important
## Plot addition

To ensure that viewers are appropriately aware of the assumption, the plot needs to have a summary somewhere of how $\tau^2$ has changed due to the inclusion of the new studies (and if possible, it's uncertainty).
:::

Another idea, that has less strong assumptions across the plot space is to consider the 'average' of the new studies, without taking into account the between-study heterogeneity, as the visual 'contribution'. As mentioned earlier, using random-weightings requires knowledge of $\tau^2$, but this isn't possible without knowing the exact 'layout' of the multiple new studies. If we remove the $\tau^2$ term, then things simplify.

Let's assume the visual contribution of the new studies (or 'average') that would be plotted (in addition to all the new studies) would be calculated as

$$\hat\theta_{\bar{new}} = \frac{\sum^m_{j=1}{\frac{\hat\theta_{nj}}{var(\hat\theta_{nj})}}}{\sum^m_{j=1}{\frac{1}{var(\hat\theta_{nj})}}}$$ $$var(\hat\theta_{\bar{new}}) = 1/\sum^m_{j=1}\frac{1}{var(\hat\theta_{nj})}$$ $m$ is the number of new studies and $\hat\theta_{nj}$ is the jth new study where $j = 1, ..., m$.

This new 'average' study is then added to the meta-analysis in the same manner as when we were only considering one new study. How much of a difference does this make?

```{r, echo = FALSE}
# Create new average study as above
MAdata_bin_new_multiple$wi <- 1/MAdata_bin_new_multiple$vi
new_avg_est <- sum(MAdata_bin_new_multiple$wi * MAdata_bin_new_multiple$yi) / sum(MAdata_bin_new_multiple$wi)
new_avg_weight <- sum(MAdata_bin_new_multiple$wi)

# Updated meta-analysis taking new studies separately
updated_meta <- metafor::rma(yi = c(MAdata_bin$yi, MAdata_bin_new_multiple$yi), sei = c(MAdata_bin$sei, MAdata_bin_new_multiple$sei), method = "DL", measure = "RR")
updated_meta

# Updated meta-analysis taking new study as one combined (weighted)
updated_meta_est <- metafor::rma(yi = c(MAdata_bin$yi, new_avg_est), vi = c(MAdata_bin$vi, 1/new_avg_weight), method = "DL", measure = "RR")
updated_meta_est
```

When including the additional studies as they are, we get a (log)RR of `{r} round(updated_meta$beta, 3)` with 95% CI: `{r} round(updated_meta$ci.lb, 3)` to `{r} round(updated_meta$ci.ub, 3)`. Whereas when we make the additional studies into one new one based on inverse-variance weighting (without accounting for between-study heterogeneity), then the (log)RR is `{r} updated_meta_est$beta` with 95% CI: `{r} round(updated_meta_est$ci.lb, 3)` to `{r} round(updated_meta_est$ci.ub, 3)`. Even though the estimates of between-study heterogeneity were similar (`{r} round(updated_meta$tau2, 3)` vs `{r} round(updated_meta_est$tau2, 3)`), the confidence bound move from being statistically significant to non-significant. This would lend itself to the visual contribution point to not match the true update (i.e., it wouldn't sit in the coloured contour, but the updated meta-analysis diamond would indicate significance). So even though this method would be closer to the 'truth' across the plot than previously, it risks there containing conflicting visual elements, rendering the plot not a useful aid.

Furthermore, it isn't possible to write the overall pooled effect (and associated standard error) in terms of this suggested formula for new 'average' study due to the $\tau^2$ term needing to be incorporating into each individual new study. I.e., if we treated the 'average' study as a single study (as is done when considering pixel by pixel), then

$$
var(\hat\theta) = 1/[\sum^n_{i=1}\frac{1}{var(\hat\theta_i) + \tau^2} + \frac{1}{var(\hat\theta_\bar{new}) + \tau^2}]
$$

but

$$
\frac{1}{var(\hat\theta_\bar{new})+\tau^2} = 1/[1/[\sum_{j=1}^m\frac{1}{var(\hat\theta_{nj})}] + \tau^2] \neq \sum_{j=1}^m\frac{1}{var(\hat\theta_{nj}) + \tau^2}
$$

A third option fits 'in between' the first two options. Let us visualise the new contribution as the weighted average, taking into account the between-study heterogeneity (estimated from the updated meta-analysis, $\hat\tau^2_{new}$).

$$
\hat\theta_{\bar{new}} = \frac{\sum^m_{j=1}{\frac{\hat\theta_{nj}}{var(\hat\theta_{nj}) + \tau^2_{new}}}}{\sum^m_{j=1}{\frac{1}{var(\hat\theta_{nj}) + \tau^2_{new}}}}
$$ {#eq-average-new-random}

$$
var(\hat\theta_{\bar{new}}) = 1/\sum^m_{j=1}\frac{1}{var(\hat\theta_{nj}) + \tau^2_{new}}
$$ {#eq-average-new-var-random}

Now let's consider each pixel on the plot. Each pixel is coloured according to the meta-analysis outcome which is derived from the pooled estimate and associated variance. When considering one new study (i.e., one new pixel), then we have the following equations (note that the $\tau^2$ term is excluded from the terms referring to new studies - this is because it's already incorporated in the 'average' values defined above):

$$
\hat\theta = \frac{\sum^n_{i=1}{\frac{\hat\theta_i}{var(\hat\theta_i) + \tau^2}} + \frac{\hat\theta_{new}}{var(\hat\theta_{new})}}{\sum^n_{i=1}\frac{1}{var(\hat\theta_i) + \tau^2} + \frac{1}{var(\hat\theta_{new})}}
$$ {#eq-pooled-est-random-new}

$$
var(\hat\theta) = 1/\sum^n_{i=1}{\frac{1}{var(\hat\theta_i) +\tau^2}} + \frac{1}{var(\hat\theta_{new})}
$$ {#eq-pooled-se-random-new}

However, as described earlier, $\tau^2$ is calculated from the existing and new studies, but in this scenario, whilst we're looking at one pixel, each pixel is representing multiple new studies, for which a unique $\tau^2$ can not be calculated.

In the first suggestion, $\tau^2 = \tau^2_{new}$ for each pixel. However, as described earlier, this leads to ill-defined contours. This third option will give an individual estimate of $\tau^2$ for each pixel, derived below.

In the appendix of Langan et. al., it shows how $\tau^2$ can be estimated using a method of moments approach:

$$
\hat\tau^2 = \frac{Q - n}{C}
$$

where Q is the test statistic in a standard test for heterogeneity, and C is a scaling factor \[Langan reference 23\]. Langan et. al. then show that for the combination of existing and new studies, Q and C can be written as

$$
Q = \sum^n_{i=1}w_i\hat\theta_i^2 + w_{new}\hat\theta^2_{new} - \frac{(\sum^n_{i=1}w_i\hat\theta_i + w_{new}\hat\theta_{new})^2}{\sum^n_{i=1}w_i + w_{new}}
$$

$$
C = \sum^n_{i=1}w_i + w_{new} - \frac{\sum^n_{i=1}w_i^2 + w^2_{new}}{\sum^n_{i=1}w_i + w_{new}}
$$

Let's simplify the notation and group together terms relating to the current studies:

$$
V = \sum^n_{i=1}w_i^2, X = \sum^n_{i=1}w_i\hat\theta_i, Y = \sum^n_{i=1}w_i, Z = \sum^n_{i=1}w_i\hat\theta_i^2
$$

noting here that $w_i$ is simply the inverse-variance (i.e., doesn't take into account any between-study heterogeneity, as that is what we're estimating). Then we get the following estimate of between-study heterogeneity

$$
\hat\tau^2 = \frac{Z + w_{new}\hat\theta^2_{new} - \frac{(X + w_{new}\hat\theta_{new})^2}{Y + w_{new}} - n}{Y + w_{new} - \frac{V + w^2_{new}}{Y + w_{new}}}
$$ {#eq-tau2-estimate}

where the weights of the studies (current and new) are the inverse variance, and $n$ is the number of current studies.

This is the case for one new study and gives a good approximation for the between-study heterogeneity of the meta-analysis including the new study. Unfortunately, when expanding @eq-tau2-estimate to consider multiple new studies, it can't be written it terms of $\hat\theta_{\bar{new}}$ and $w_{\bar{new}}$, as we did for the fixed effects scenario due to the squared terms. We also can't plug in the 'average' values, as those values include between-study heterogeneity - which is what is being estimated. However, varying @eq-tau2-estimate slightly, could give us a way to appropriately take into account between-study heterogeneity as we move further away from the 'average' new study that we do have data for.

For each pixel, let $\tau^2$ be estimated with the below:

$$
\hat\tau^2 = \frac{Z + \frac{\hat\theta^2_{new}}{var(\hat\theta_{new})} - \frac{(X + \frac{\hat\theta_{new}}{var(\hat\theta_{new})})^2}{Y + \frac{1}{var(\hat\theta_{new})}} - (n+m-1)}{Y + \frac{1}{var(\hat\theta_{new})} - \frac{V + \frac{1}{var(\hat\theta_{new})^2}}{Y + \frac{1}{var(\hat\theta_{new})}}} * K
$$ {#eq-tau2-estimate-adj}

where $n$ and $m$ are the number of existing and new studies respectively (and so accounting for the fact that this is actually multiple new studies rather than 1). $K$ is calculated by substituting $\hat\tau^2$ with $\tau^2_{new}$, and $\hat\theta_{new}$ and $var(\hat\theta_{new})$ representing the 'average' new effect using @eq-pooled-est-random-new and @eq-pooled-se-random-new. Then $K$ is a scaling factor that ensures that at the co-ordinates that represent the 'average' new study, the pixel will be coloured according to meta-analysis results that are already known from including the new studies (and so marries up with the summary diamonds).

Therefore, using @eq-tau2-estimate-adj essentially allows us to 'start' with the 'known' between-study heterogeneity estimate, and then adjust the between-study heterogeneity appropriately as we move further away from the 'average' new study.

```{r, echo = FALSE}
# Obtain tau2_new
updated_meta <- metafor::rma(yi = c(MAdata_bin$yi, MAdata_bin_new_multiple$yi), sei = c(MAdata_bin$sei, MAdata_bin_new_multiple$sei), method = "DL", measure = "RR")
updated_tau2 <- updated_meta$tau2

# Obtain average new study effects
MAdata_bin_new_multiple$wi <- 1/(MAdata_bin_new_multiple$vi + updated_tau2)
new_avg_est <- sum(MAdata_bin_new_multiple$wi * MAdata_bin_new_multiple$yi) / sum(MAdata_bin_new_multiple$wi)
new_avg_var <- 1/sum(MAdata_bin_new_multiple$wi)

# Calculate adjustment factor for this example
# Create V, X, Y, and Z
MAdata_bin$wi <- 1/MAdata_bin$vi
V <- sum(MAdata_bin$wi^2)
X <- sum(MAdata_bin$wi * MAdata_bin$yi)
Y <- sum(MAdata_bin$wi)
Z <- sum(MAdata_bin$wi * MAdata_bin$yi^2)
# Calculate K
K <- updated_tau2 / ((Z + (new_avg_est^2/new_avg_var) - ((X + new_avg_est/new_avg_var)^2/(Y + 1/new_avg_var)) - 8)/(Y + 1/new_avg_var - ((V + 1/new_avg_var^2)/(Y + 1/new_avg_var))))
```

In our example, $\tau^2_{new}$ is estimated as being `{r} round(updated_tau2, 3)` and the weighted average values (using this estimated between-study heterogeneity) $\theta_{\bar{new}}$ and $var(\hat\theta_{\bar{new}})$ are estimated to be `{r} round(new_avg_est, 3)` and `{r} round(new_avg_var, 3)`. Plugging these into equation @eq-tau2-estimate-adj, gives us an scaling factor of `{r} round(K, 3)`.

##### Example plot

One notices that with the example plot below, it behaves very differently to when considering one new study. This is because for multiple studies under a random-effects framework, the pixels are actually representing something slightly different - each pixel is already partially taking into account the estimated between-study heterogeneity. This is taken into account by removing the $\tau^2$ term from the new study elements of @eq-pooled-est-random-new and @eq-pooled-se-random-new, and it's this removal of $\tau^2$ that causes there to be differently coloured pixels compared to if only including one study. Essentially, when the viewer is considering other pixels away from the 'average' found from the new studies, they need to consider whether such a random-weighted average would likely exist (i.e., the colour represents the meta-analysis result *assuming* that the average of $m$ new studies, incorporating the estimated between-study heterogeneity, could sit at those pixel coordinates). Therefore, in this setting, the plot probably needs a warning to be careful when extracting far beyond the new data.

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 50)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

Because the between-study heterogeneity depends on all the studies, the contours technically depend on the new studies that were found (whereas the contours are 'independent' of the new studies in the other plots). But how different would they be? Let's see the effect of a different set of three studies that show an opposite effect to previously.

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = -MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 50)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

The resulting plot is actually fairly similar, suggesting that it might not be as dangerous to extrapolate beyond the 'average' point as first anticipated.

Lets see what happens when we have three new studies who, on their own, would have led to a non-significant result under a random-effects framework.

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = c(log(0.26), log(0.5), log(0.35)), seSSnew = c(0.05, 0.05, 0.12),
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 50)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

Two interesting things to note here. Firstly, whilst each of the new studies individually wouldn't have led to a significant result, having three in this extreme area does lead to statistical significance (likely because having multiple studies in this extreme area (rather than one) possibly reduces the between-study heterogeneity). Secondly, the weighted 'average' new study, is nowhere near the new studies on this plot - illustrating that under this interpretation of the pixels, pixels that are far away from the general 'crowd' of studies are unlikely to exist in practicality (especially the areas of small standard error, due to the presence of the between-study heterogeneity term).

::: callout-important
## Plot additions

-   Dotted lines between the individual new studies and the overall 'average' new study (making a spider-like shape).

-   Choose a different shape for the 'average' new study, as using a diamond confuses people that there may be something associated with the width of the diamond (as is the case for the pooled effects).
:::

## Compare each case

```{r, echo = FALSE, include = FALSE}
result_fixed_single <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'fixed', outcome = 'RR', zero = 1, sig_level = 0.1)
result_random_single <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 50)
result_fixed_mult <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'fixed', outcome = 'RR', zero = 1, sig_level = 0.1)
result_random_mult <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 100)
```

```{r, echo = FALSE, fig.width = 10}
result_fixed_single$threshold_plot + ggtitle("Fixed-effects with single new study")

result_random_single$threshold_plot + ggtitle("Random-effects with single new study")

result_fixed_mult$threshold_plot + ggtitle("Fixed-effects with multiple new studies")

result_random_mult$threshold_plot + ggtitle("Random-effects with multiple new studies")
```
