---
title: "Interpretation Tool"
format: docx
editor: visual
---

This document is to go over the details of the interpretation tool part of the update framework, which is based on the work by Langan et. al.

```{r setup, include = FALSE}

# Load function and example datasets
source("InterpretationTool.R")

# Remove progress bars when rendering
if (!interactive()) {
  progressr::handlers("void")
}

### Test Data antibiotics vs. control for the common cold to alleviate symptoms by 7 days ###
raw_data_bin <- data.frame(StudyID = c(1, 2, 3, 4, 5, 6), Study = c("Herne_1980", "Hoaglund_1950", "Kaiser_1996", "Lexomboon_1971", "McKerrow_1961", "Taylor_1977"),
                   R.1 = c(7, 39, 97, 8, 5, 12), N.1 = c(7+39, 39+115, 97+49, 8+166, 5+10, 12+117), T.1 = rep("Treatment", 6),
                   R.2 = c(10, 51, 94, 4, 8, 3), N.2 = c(10+12, 51+104, 94+48, 4+83, 8+10, 3+56), T.2 = rep("Control", 6))
raw_data_new <- data.frame(StudyID = c(7), Study = c("Fake_1"),
                           R.1 = c(63), N.1 = c(63+252), T.1 = rep("Treatment", 1),
                           R.2 = c(77), N.2 = c(77+105), T.2 = rep("Control", 1))
raw_data_new_multiple <- data.frame(StudyID = c(7, 8, 9), Study = c("Fake_1", "Fake_2", "Fake_3"),
                           R.1 = c(63, 12, 45), N.1 = c(63+252, 12+34, 45+60), T.1 = rep("Treatment", 3),
                           R.2 = c(77, 21, 58), N.2 = c(77+105, 21+25, 58+42), T.2 = rep("Control", 1))

### Obtain study effects and standard errors #
MAdata_bin <- metafor::escalc(measure = 'RR', ai = R.1, bi = N.1-R.1, ci = R.2, di = N.2-R.2, data = raw_data_bin)   # gives ES (effect estimate) and seES (sampling variances) on logOR scale for binary data
MAdata_bin_new <- metafor::escalc(measure = 'RR', ai = R.1, bi = N.1-R.1, ci = R.2, di = N.2-R.2, data = raw_data_new)
MAdata_bin_new_multiple <- metafor::escalc(measure = 'RR', ai = R.1, bi = N.1-R.1, ci = R.2, di = N.2-R.2, data = raw_data_new_multiple)
MAdata_bin$sei <- sqrt(MAdata_bin$vi)  # Calculate standard errors
MAdata_bin_new$sei <- sqrt(MAdata_bin_new$vi)
MAdata_bin_new_multiple$sei <- sqrt(MAdata_bin_new_multiple$vi)
```

# Derivation of contours

This section will go over the derivation of forming the contours of the plot for different situations.

## Fixed-effects

We are working within the fixed-effects framework initially, where the between-study heterogeneity, denoted by $\tau^2$, is equal to zero.

### Statistical significance

Lets assume we are interested in changing the statistical significance (i.e., the p-value).

#### One new study

Let's start with assuming that there is only one new study that will be added to the meta-analysis.

The 95% confidence interval (CI) is denoted by

$$\hat{\theta} \pm \Phi^{-1} (\frac{\alpha}{2}) \sqrt{var(\hat\theta)}$$ {#eq-ci}

where $\hat\theta$ is the pooled effect estimate, $\alpha$ is the type I error rate (normally set at 0.05), $\Phi^{-1}$ is the inverse cumulative normal distribution with mean 0 and variance 1 (which ensures that $\Phi^{-1}(\alpha/2)$ maps an $\alpha$ value of 0.05 to the value of 1.96). $var(\hat\theta)$ is the variance of the pooled effect estimate.

To create a threshold for statistical significance (specified by $\alpha$), we can specify that the lower or upper bound of the 95% CI needs to touch the null value, denoted by $\theta_0$.

Let's assume that there is a new study with (unknown) effect estimate $\hat\theta_{new}$ , variance $var(\hat\theta_{new})$, and weight $w_{new}$. Then once this new study is added to the meta-analysis, the updated results will be as follows:

$$
\hat\theta = \frac{\sum^{n}_{i=1}{w_i\hat\theta_i} + w_{new}\hat\theta_{new}}{\sum^n_{i=1}{w_i} + w_{new}}
$$ {#eq-updated-estimate}

$$var(\hat\theta) = \frac{1}{\sum^n_{i=1}{w_i} + w_{new}}$$ {#eq-updated-variance}

where $n$ is the number of studies in the original meta-analysis.

By substituting @eq-updated-estimate and @eq-updated-variance into @eq-ci, and specifying that the upper or lower bound equals $\theta_0$, we can get an expression for $\theta_{new}$ in terms of other quantities:

$$
\hat\theta_{new} = \frac{1}{w_{new}}\{\theta_0[\sum^n_{i=1}{w_i} + w_{new}] - \sum^n_{i=1}{w_i \hat\theta_i} \pm \Phi^{-1}(\frac{\alpha}{2}) \sqrt{\sum^n_{i=1}{w_i} + w_{new}}\}
$$ {#eq-rearranged-theta-new}

In a fixed-effects meta-analysis, each weight is simply calculated as

$$
w_i = \frac{1}{var(\hat\theta_i)}
$$ {#eq-fixed-weight}

Therefore, by substituting @eq-fixed-weight into @eq-rearranged-theta-new, we then get an equation for the contour threshold on the estimate-variance plane.

##### Example plot

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'fixed', outcome = 'RR', zero = 1)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

#### Two new studies

When expanding from one new study to two new studies, @eq-updated-estimate and @eq-updated-variance become

$$
\hat\theta = \frac{\sum^{n}_{i=1}{w_i\hat\theta_i} + w_{n1}\hat\theta_{n1} + w_{n2}\hat\theta_{n2}}{\sum^n_{i=1}{w_i} + w_{n1} + w_{n2}}
$$ {#eq-meta-theta-2studies}

$$var(\hat\theta) = \frac{1}{\sum^n_{i=1}{w_i} + w_{n1} + w_{n2}}$$ {#eq-meta-variance-2studies}

where $\hat\theta_{nj}$ and $w_{nj}$ are the estimates and weights of new studies respectively, where $j = 1,2$.

Let

$$w_{\bar{new}} = w_{n1} + w_{n2}$$ {#eq-weight-2studies}

$$\hat\theta_{\bar{new}} = \frac{w_{n1}\hat\theta_{n1} + w_{n2}\hat\theta_{n2}}{w_{n1} + w_{n2}}$$ {#eq-theta-2studies} Substituting @eq-weight-2studies and @eq-theta-2studies into @eq-meta-theta-2studies and @eq-meta-variance-2studies gives us the following:

$$
\hat\theta = \frac{\sum^{n}_{i=1}{w_i\hat\theta_i} + w_{\bar{new}}\hat\theta_{\bar{new}}}{\sum^n_{i=1}{w_i} + w_{\bar{new}}}
$$

$$
var(\hat\theta) = \frac{1}{\sum^n_{i=1}{w_i} + w_{\bar{new}}}
$$

This is simply the same equations as equations @eq-updated-estimate and @eq-updated-variance. Therefore, everything else follows through as before, but with $w_{\bar{new}}$ and $\hat\theta_{\bar{new}}$ in place of $w_{new}$ and $\hat\theta_{new}$.

#### Multiple new studies {#sec-multiple-new-studies-fixed}

Where we are calculating the contour to $m$ new studies, the equations follow as if it was one new study, but $w_{new}$ and $\hat\theta_{new}$ are defined as

$$
w_{\bar{new}} = \sum^m_{j=1}{w_{nj}}
$$

$$
\hat\theta_{\bar{new}} = \frac{\sum^m_{j=1}{w_{nj}\hat\theta_{nj}}}{\sum^m_{j=1}{w_{nj}}}
$$

where $\hat\theta_{nj}$ and $w_{nj}$ are the estimates and weights of the new studies respectively, for $j = 1,2,...,m$.

##### Example plot

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'fixed', outcome = 'RR', zero = 1, sig_level = 0.0005)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

### Specific confidence interval bounds

Let's suppose that instead of wanting a statistically significant pooled estimate, that the researchers are looking for the 95% CI bounds to be at a certain value (likely of clinical importance).

With statistical significance, the contour equations are calculated from stating that the upper or lower bounds must touch the null value, $\theta_0$. In this scenario, the contour equations would be calculated from the following thresholds:

$$
\theta_{lower} = \hat{\theta} - \Phi^{-1} (\frac{\alpha}{2}) \sqrt{var(\hat\theta)}
$$ {#eq-lower-bound}

$$
\theta_{upper} = \hat{\theta} + \Phi^{-1} (\frac{\alpha}{2}) \sqrt{var(\hat\theta)}
$$ {#eq-upper-bound}

where $\theta_{lower}$ and $\theta_{upper}$ are the user-defined lower and upper 95% CI bounds of interest.

Substituting @eq-updated-estimate and @eq-updated-variance into @eq-lower-bound and @eq-upper-bound, and rearranging gives the following expression(s) for $\hat\theta_{new}$.

$$
\hat\theta_{new} = \frac{1}{w_{new}}\{\theta_{lower}[\sum^n_{i=1}{w_i} + w_{new}] - \sum^n_{i=1}{w_i \hat\theta_i} + \Phi^{-1}(\frac{\alpha}{2}) \sqrt{\sum^n_{i=1}{w_i} + w_{new}}\}
$$ $$
\hat\theta_{new} = \frac{1}{w_{new}}\{\theta_{upper}[\sum^n_{i=1}{w_i} + w_{new}] - \sum^n_{i=1}{w_i \hat\theta_i} - \Phi^{-1}(\frac{\alpha}{2}) \sqrt{\sum^n_{i=1}{w_i} + w_{new}}\}
$$

Whilst in the fixed-effects framework, the extensions to multiple new studies is the same here as for statistical significance.

##### Example plot

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'fixed', outcome = 'RR', upper = 0.9)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

## Random-effects

When working in the random-effects framework things get more complex as $\tau^2 \ne 0$ and is estimated from the studies in the meta-analysis, including the new studies.

Let's start again with only considering one new study and aiming for a meta-analysis with statistical significance at the $\alpha = 0.05$ level.

Under a random-effects framework, the weight for each study (including new studies) is instead calculated as

$$
w_i = \frac{1}{var(\hat\theta_i) + \tau^2_{new}}
$$ {#eq-random-weight}

where $\tau^2_{new}$ is the estimated between-study heterogeneity when included the new studies into the meta-analysis. Unfortunately, we can't simply plug the random-effects equation for weights into the contour equations derived for fixed effects (@eq-rearranged-theta-new) as $\tau^2_{new}$ depends on $\hat\theta_{new}$. Furthermore, we can't simply plug in an equation for $\tau^2_{new}$ and rearrange as calculating an estimate for between-study heterogeneity is too complicated to do by hand, and so researchers often use established estimators instead, such as the DerSimonian-Laird estimator or Restricted Maximum Likelihood. (Note the further challenge, that these commonly used estimators wouldn't allow us to incorporate the uncertainty in estimating between-study heterogeneity).

As such, an equation to form the contour thresholds cannot be derived as was done for the fixed-effects framework. Instead, as per Langan et. al., the standard error-estimate plane is broken down into 'pixels', where $\tau^2_{new}$ is estimated at each pixel (calculated from the original studies as well as the new study represented by the pixel). For each pixel, $\tau^2_{new}$, $\hat\theta_{new}$, and $var(\hat\theta_{new})$ are plugged into the meta-analysis equations ( @eq-updated-estimate and @eq-updated-variance), but using the weights calculated in @eq-random-weight, and noted whether or not the threshold was crossed (i.e. whether 95% CI was one side of null effect). The result at each pixel is coloured according to whether the threshold was met. This easily extends to the clinical bound thresholds as well as the statistical significance.

##### Example plot

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 100)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

### Multiple new studies {#multiple-new-studies-1}

Under the random-effects model, we have to take the estimate-variance plane 'pixel-by-pixel'. When there was one new study, this was simple as the pixel had one set of possible values 'attached' to it that were easily obtained. When considering multiple new studies, the idea of using the 'average' of the new studies (as was done for fixed-effects) doesn't work for multiple reasons. Firstly, for each pixel, it is not possible to know what new studies created the 'average' represented by the respective point. Without knowing the new studies, the between-study heterogeneity can't be estimated. Secondly, different combinations of new studies could lead to the same 'average' effect size and standard error. It therefore follows that as there is no unique combination, there could be multiple different estimates of the between-study heterogeneity represented by a single pixel. To move forward, a compromise or assumption has to be made.

**Let's assume a constant** **between-study heterogeneity** **across all pixels in the plane.**

The estimated between-study heterogeneity from the updated meta-analysis, $\tau^2_{new}$, is proposed as the common between-study heterogeneity across all pixels. Whilst this may be a strong assumption as you get to more 'extreme' pixels/values, the new studies 'average' would at least sit in the correct section of the plot (as per the fixed-effects).

::: callout-important
## Plot addition

To ensure that viewers are appropriately aware of the assumption, the plot needs to have a summary somewhere of how $\tau^2$ has changed due to the inclusion of the new studies (and if possible, it's uncertainty).
:::

The new studies 'average' that would be plotted (in addition to all the new studies) would be calculated as

$$\hat\theta_{\bar{new}} = \frac{\sum^m_{j=1}{\frac{\hat\theta_{nj}}{var(\hat\theta_{nj}) + \tau^2_{new}}}}{\sum^m_{j=1}{\frac{1}{var(\hat\theta_{nj}) + \tau^2_{new}}}}$$ {#eq-average-new-random} $$var(\hat\theta_{\bar{new}}) = 1/\sum^m_{j=1}\frac{1}{var(\hat\theta_{nj}) + \tau^2_{new}}$$ {#eq-average-new-var-random} where $m$ is the number of new studies and $\hat\theta_{nj}$ is the jth new study where $j = 1, ..., m$.

As we're interested in how the 'average' new study (or contribution) impacts the updated meta-analysis, we need to consider the pooled estimate (and associated standard error/variance) in terms of the 'average' new study. The following equations satisfy this:

$$
\hat\theta = \sum^n_{i=1}{\frac{\hat\theta_i}{var(\hat\theta_i) + \tau^2_{new}}} + \frac{\hat\theta_{\bar{new}}}{var(\hat\theta_{\bar{new}})}
$$ {#eq-pooled-est-random-multiple}

$$
var(\hat\theta) = 1/\sum^n_{i=1}{\frac{1}{var(\hat\theta_i) +\tau^2_{new}}} + \frac{1}{var(\hat\theta_{new})}
$$ {#eq-pooled-se-random-multiple}

Therefore, to build contours that show the effect of an 'average' new study, pixel by pixel, @eq-pooled-est-random-multiple and @eq-pooled-se-random-multiple are used, where $\hat\theta_{new}$ and $var(\hat\theta_{new})$ are substituted by the pixel values, and $\tau^2_{new}$ is the value from the updated meta-analysis. As per the scenario with one new study, each pixel is coloured according to whether the specified threshold was met. Note that $\tau^2_{new}$ is already incorporated into the new contributions to the pooled effect through @eq-average-new-random and @eq-average-new-var-random, which is why $\tau^2_{new}$ isn't needed in the latter elements of the equations above. This isn't needed for the fixed-effects model when considering multiple new studies due to how the weights are calculated (i.e., weight is just the inverse of the variance, which is different for each study).

A potential extension, depending on time and resources, could be to estimate $\tau^2_{new}$ at each pixel by simulating $m$ new studies. This decision depends on the balance between computational resource and relaxing the assumption made regarding a fixed $\tau^2_{new}$ for all pixels. In the current scenario (i.e., being a visual aid), I feel that the assumption above is satisfactory - but this may change depending on how the plot's use is extended.

##### Example plot

```{r, echo = FALSE, include = FALSE}
result <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.05)
```

```{r, echo = FALSE, fig.width = 10}
result$threshold_result

result$threshold_plot
```

::: callout-important
## Plot additions

-   Dotted lines between the individual new studies and the overall 'average' new study (making a spider-like shape).

-   Choose a different shape for the 'average' new study, as using a diamond confuses people that there may be something associated with the width of the diamond (as is the case for the pooled effects).
:::

## Interesting comparisons

### Fixed vs Random (One new study)

```{r, echo = FALSE, include = FALSE}
result_fixed <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'fixed', outcome = 'RR', zero = 1, sig_level = 0.1)
result_random <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 50)
```

```{r, echo = FALSE, fig.width = 10}
# | fig-cap: "One new study"
# | fig-subcap:
# |   - "Fixed-effects"
# |   - "Random-effects"
# | layout-nrow: 2

result_fixed$threshold_plot

result_random$threshold_plot
```

### Fixed vs Random (Multiple studies)

```{r, echo = FALSE, include = FALSE}
result_fixed <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'fixed', outcome = 'RR', zero = 1, sig_level = 0.05)
result_random <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.05, contour_points = 100)
```

```{r, echo = FALSE, fig.width = 10}
# | fig-cap: "Multiple new studies"
# | fig-subcap:
# |   - "Fixed-effects"
# |   - "Random-effects"
# | layout-nrow: 2

result_fixed$threshold_plot

result_random$threshold_plot
```

### One vs multiple new studies (Random-effects)

```{r, echo = FALSE, include = FALSE}
result_one <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new$yi, seSSnew = MAdata_bin_new$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 100)
result_multiple <- InterpretationThreshold(SS = MAdata_bin$yi, seSS = MAdata_bin$sei,
  SSnew = MAdata_bin_new_multiple$yi, seSSnew = MAdata_bin_new_multiple$sei,
  method = 'random', outcome = 'RR', zero = 1, sig_level = 0.1, contour_points = 100)
```

```{r, echo = FALSE, fig.width = 10}
# | fig-cap: "Random-effects"
# | fig-subcap:
# |   - "One new study"
# |   - "Multiple new studies"
# | layout-nrow: 2

result_one$threshold_plot

result_multiple$threshold_plot
```

## Addressing assumption made for multiple studies in random-effects model

Currently, due to the assumption about the between-study heterogeneity being fixed at each pixel to the updated meta-analysis value, some areas of the plot are 'far-fetched'. For example, in the final example shown above, the white region in the top left is white with one new study, because the between-study heterogeneity at that point would become so large, that the pooled effect wouldn't be significant. However, when considering multiple studies, because I've fixed the between-study heterogeneity (to a value that is potentially unrealistic), the weightings aren't as large, and so a 'significant' effect is appearing.

***Is there a way to improve this assumption?***

We've already discussed how it is not possible to get the exact estimates of between-study heterogeneity for each pixel when considering multiple studies. However, is there a way to account for the likely change in heterogeneity as we move further away from the 'average' new study point (for which the heterogeneity is currently fixed)?

In the appendix of Langan et. al., it shows how $\tau^2$ can be estimated using a method of moments approach:

$$
\hat\tau^2 = \frac{Q - n}{C}
$$

where Q is the test statistic in a standard test for heterogeneity, and C is a scaling factor \[Langan reference 23\]. Langan et. al. then show that for the combination of existing and new studies, Q and C can be written as

$$
Q = \sum^n_{i=1}w_i\hat\theta_i^2 + w_{new}\hat\theta^2_{new} - \frac{(\sum^n_{i=1}w_i\hat\theta_i + w_{new}\hat\theta_{new})^2}{\sum^n_{i=1}w_i + w_{new}}
$$

$$
C = \sum^n_{i=1}w_i + w_{new} - \frac{\sum^n_{i=1}w_i^2 + w^2_{new}}{\sum^n_{i=1}w_i + w_{new}}
$$

Let's simplify the notation and group together terms relating to the current studies:

$$
V = \sum^n_{i=1}w_i^2, X = \sum^n_{i=1}w_i\hat\theta_i, Y = \sum^n_{i=1}w_i, Z = \sum^n_{i=1}w_i\hat\theta_i^2
$$

Then we get the following estimate of between-study heterogeneity

$$
\hat\tau^2 = \frac{Z + w_{new}\hat\theta^2_{new} - \frac{(X + w_{new}\hat\theta_{new})^2}{Y + w_{new}} - n}{Y + w_{new} - \frac{V + w^2_{new}}{Y + w_{new}}}
$$ {#eq-tau2-estimate}

where the weights of the studies (current and new) are the inverse variance, and $n$ is the number of current studies.

This is the case for one new study and gives a good approximation for the between-study heterogeneity of the meta-analysis included the new study. How can this be extended to multiple new studies?

The terms V, X, Y, and Z all stay the same as they are calculated from the existing studies. Currently, the plot is looking at each pixel as if it is representing $\hat\theta_{\bar{new}}$ and $w_{\bar{new}}$, and assuming $\tau^2_{new}$ to be constant across all pixels. The ideal scenario would be to extend @eq-tau2-estimate to multiple new studies, and rearrange to be in terms of $\hat\theta_{\bar{new}}$ and $w_{\bar{new}}$ (as we did with the fixed-effects scenario) to then get an estimate of $\tau^2$ at each pixel, however, this is not possible (due to the squared terms). Furthermore, currently each pixel represents the weight under the random-effects model (i.e., incorporates $\tau^2$, whereas in @eq-tau2-estimate, the weights are simply the inverse variances.

There are two things that need sorting:

1.  How to transform $w_{new}$ such that is represents $m$ new studies, rather than 1.

2.  How do we represent new studies in the plot.

Regarding the second point, as the contours are going to rely on $\tau^2$, and that is going to be estimated from @eq-tau2-estimate, to keep things visually cohesive, the 'average' new contribution is going to have to be a simple weighted average (i.e., not take $\tau^2$ into account). Note that this doesn't mean $\tau^2$ will be ignored when carrying out the meta-analysis at each pixel.

To help with point 1, I first want to see how the estimate of $\tau^2$ differs between adding three new studies and adding one new study that is the combined (raw) results.

```{r, echo = FALSE}
# Create combined new study
Fake.R1 <- sum(MAdata_bin_new_multiple$R.1)
Fake.N1 <- sum(MAdata_bin_new_multiple$N.1)
Fake.R2 <- sum(MAdata_bin_new_multiple$R.2)
Fake.N2 <- sum(MAdata_bin_new_multiple$N.2)
Fake.y <- as.numeric(metafor::escalc(measure = 'RR', ai = Fake.R1, n1i = Fake.N1, ci = Fake.R2, n2i = Fake.N2)[1])
Fake.v <- as.numeric(metafor::escalc(measure = 'RR', ai = Fake.R1, n1i = Fake.N1, ci = Fake.R2, n2i = Fake.N2)[2])

# Create V, X, Y, and Z
MAdata_bin$wi <- 1/MAdata_bin$vi
V <- sum(MAdata_bin$wi^2)
X <- sum(MAdata_bin$wi * MAdata_bin$yi)
Y <- sum(MAdata_bin$wi)
Z <- sum(MAdata_bin$wi * MAdata_bin$yi^2)

# Test estimate for one new study
MAdata_bin_new$wi <- 1/MAdata_bin_new$vi
tau2_one <- as.numeric((Z + MAdata_bin_new$wi * MAdata_bin_new$yi^2 - (X + MAdata_bin_new$wi * MAdata_bin_new$yi)^2/(Y + MAdata_bin_new$wi) - (6)) / (Y + MAdata_bin_new$wi - (V + MAdata_bin_new$wi^2)/(Y + MAdata_bin_new$wi)))
updated_meta <- metafor::rma(yi = c(MAdata_bin$yi, MAdata_bin_new$yi), sei = c(MAdata_bin$sei, MAdata_bin_new$sei), method = "DL", measure = "RR")
updated_tau2 <- updated_meta$tau2

# tau2 estimate for three individual studies
MAdata_bin_new_multiple$wi <- 1/MAdata_bin_new_multiple$vi
tau2_ind <- (Z + sum(MAdata_bin_new_multiple$wi * MAdata_bin_new_multiple$yi^2) - (X + sum(MAdata_bin_new_multiple$wi * MAdata_bin_new_multiple$yi))^2/(Y + sum(MAdata_bin_new_multiple$wi)) - (6+2)) / (Y + sum(MAdata_bin_new_multiple$wi) - (V + sum(MAdata_bin_new_multiple$wi^2))/(Y + sum(MAdata_bin_new_multiple$wi)))

#tau2 estimate as if one study
Fake.w <- 1/Fake.v
tau2_comb <- (Z + Fake.w * Fake.y^2 - (X + Fake.w * Fake.y)^2/(Y + Fake.w) - (6)) / (Y + Fake.w - (V + Fake.w^2)/(Y + Fake.w))
# does it improve when we pretend its 3 new studies
tau2_comb_adj1 <- (Z + Fake.w * Fake.y^2 - (X + Fake.w * Fake.y)^2/(Y + Fake.w) - (6+2)) / (Y + Fake.w - (V + Fake.w^2)/(Y + Fake.w))

# Use weighted average instead of combining raw numbers?
new_avg_est <- sum(MAdata_bin_new_multiple$wi * MAdata_bin_new_multiple$yi) / sum(MAdata_bin_new_multiple$wi)
new_avg_weight <- sum(MAdata_bin_new_multiple$wi)
tau2_comb_adj2 <- (Z + new_avg_weight * new_avg_est^2 - (X + new_avg_weight * new_avg_est)^2/(Y + new_avg_weight) - (6)) / (Y + new_avg_weight - (V + new_avg_weight^2)/(Y + new_avg_weight))

# true tau2 from adding new studies
updated_meta <- metafor::rma(yi = c(MAdata_bin$yi, MAdata_bin_new_multiple$yi), sei = c(MAdata_bin$sei, MAdata_bin_new_multiple$sei), method = "DL", measure = "RR")
updated_tau2 <- updated_meta$tau2
```

When adding our example three new studies, the estimated $\tau^2$ from metafor is `{r} round(updated_tau2, 3)`. For estimating $\tau^2$ using the method of moments equations, when the new studies were considered individually, $\tau^2$ was estimated to be `{r} round(tau2_ind, 3)`, which is a very close estimate to the estimated value from metafor. However, when the new studies were combined into one individual study, $\tau^2$ was estimated to be `{r} round(tau2_comb, 3)`, which is slightly off.

The equation for estimating $\tau^2$ has a term for the number of studies in the meta-analysis. Naturally, when we add more new studies, this term increases by $m-1$, where $m$ is the number of new studies. When taking the combined study, but fixing the aforementioned term to reflect the 'true' number of new studies, we get an estimate of `{r} round(tau2_comb_adj1, 3)`, which is an improvement, but still not ideal.

Instead of combining the new studies using the raw data, let's combine then by their weights - creating their own pooled effect. As we don't have an estimate of $\tau^2$, we'll just be using the inverse-variance as weights (as for the fixed-effects, and @eq-tau2-estimate). In this case, the estimate gives `{r} round(tau2_comb_adj2, 3)` (note that this doesn't include the adjustment regarding number of studies that was explored above). This is a pretty good estimate, and one I'd be happy to take to get better contours for the plot!

So, at each pixel, we estimate $\tau^2_{new}$ using @eq-tau2-estimate, where each pixel represents the weighted average of the new studies (where weights are simply inverse-variances). We then plug the estimate into the meta-analysis command, and colour according to whether the pooled result gives the desired effect.
